### 2024-03-28 paper/paper.qmd
model result

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Mary Cheng
thanks: "Code and data are available at: https://github.com/marycx/us_poverty_analysis_2019.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(marginaleffects)
library(plotly)
library(tibble)
library(margins)
library(kableExtra)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

```{r}

```

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{mortgage}_i + \beta_2 \times \mbox{income}_i + \beta_3 \times \mbox{age}_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-modelresults
#| tbl-cap: "Explanatory model Poverty Prediction (n = 1000)"

poverty_prediction_model <-
  readRDS(file = here::here("models/poverty_prediction_model.rds"))

modelsummary(
  list(
    "In poverty" = poverty_prediction_model
  ),
  statistic = "mad"
  )
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Model details {#sec-model-details}

## Posterior predictive check
In @fig-post_dist, we implement a posterior distribution. This compares the poverty status of people in reality with the prediction results from the posterior distribution from our logistic regression model. It can be seen that the posterior distribution fits well with the actual data. It suggests that the posterior is able to generate simulated data that closely resembles the actual data [@bayes_post], because the model accurately captures the observed data patterns. This is good because it indicates that our logistic regression model is a good representation of the actual poverty status in the 2019 poverty data from United States Census Bureau.

```{r}
#| label: fig-post_dist
#| fig-cap: Posterior distribution for logistic regression model 
#| echo: false
#| warning: false
#| message: false
#|

pp_check(poverty_prediction_model) +
  theme(legend.position = "bottom")
```

@fig-post_prior compares the posterior with the prior. This compares the prior distribution of parameters with the posterior distribution of parameters in our logistic regression model. We can see that half of the model parameters do not change after data are taken into account, while some parameters shift slightly. This shows that the observed data partially matches with our initial belief and expectation about the poverty status of people in the United States in 2019. We can see that for people with income level at "below 10k", the posterior distribution greatly shifts from its prior after we input observed data. This is likely suggesting that the observed data for "below 10k" strongly contradicts our initial belief. This suggests that the majority of people in US who earns less than a total amount of 10k per household annually is in poverty in 2019.

```{r}
#| label: fig-post_prior
#| fig-cap: Comparing the posterior with the prior 
#| echo: false
#| warning: false
#| message: false

#### Read data ####
analysis_data <- read_parquet(here::here("data/analysis_data/cleaned_poverty_data.parquet"))

# Convert variables to factors
analysis_data$mortgage <- factor(analysis_data$mortgage)
analysis_data$income <- factor(analysis_data$income)
analysis_data$age <- factor(analysis_data$age)

# Create poverty_status variable in binary form
analysis_data$poverty_status <- ifelse(analysis_data$poverty_status == "In poverty", 1, 0)

# Model 1 for n = 1000
set.seed(215)

poverty_data_reduced <- 
  analysis_data |> 
  slice_sample(n = 1000)

posterior_vs_prior(poverty_prediction_model) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Markov chain Monte Carlo Convergence Check

@fig-trace1, @fig-trace2, @fig-trace3 are the trace plots of the model. It tells us if there is existence of signs that the our model runs into issues. We observe lines in all the trace plots are horizontal and oscillating, and have overlaps between the chains. This suggests that there is nothing strange in this trace plot.

@fig-rhat is the Rhat plot of the model. It compares the variability within each chain to the variability between chains in MCMC. We can observe that our Rhat plot are all close to 1, and no more than 1.05. This is a good sign because it suggests that the MCMC algorithm has reached convergence for our model.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace1
#| fig-cap: "Trace plot of intercept and mortgage"
#| fig-subcap: ["Trace plot of Intercept", "Trace plot of owner without mortgage", "Trace plot of renter"]
#| layout-ncol: 2

plot(poverty_prediction_model, "trace", "(Intercept)")
plot(poverty_prediction_model, "trace", "mortgageOwner without mortgage")
plot(poverty_prediction_model, "trace", "mortgageRenter")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace2
#| fig-cap: "Trace plot of income"
#| fig-subcap: ["Trace plot of income 10k-50k", "Trace plot of income 150k-200k", "Trace plot of income 200k-250k", "Trace plot of income 50k-100k", "Trace plot of income above 250k", "Trace plot of income below 10k"]
#| layout-ncol: 2

plot(poverty_prediction_model, "trace", "income10k-50k")
plot(poverty_prediction_model, "trace", "income150k-200k")
plot(poverty_prediction_model, "trace", "income200k-250k")
plot(poverty_prediction_model, "trace", "income50k-100k")
plot(poverty_prediction_model, "trace", "incomeabove 250k")
plot(poverty_prediction_model, "trace", "incomebelow 10k")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace3
#| fig-cap: "Trace plot of employ"
#| fig-subcap: ["Trace plot of age 35-44", "Trace plot of age 45-54", "Trace plot of age 55-64", "Trace plot of age above 65"]
#| layout-ncol: 2

plot(poverty_prediction_model, "trace", "age35-44")
plot(poverty_prediction_model, "trace", "age45-54")
plot(poverty_prediction_model, "trace", "age55-64")
plot(poverty_prediction_model, "trace", "ageabove 65")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-rhat
#| fig-cap: "Rhat plot"

plot(poverty_prediction_model, "rhat")
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of support for Biden"

modelplot(poverty_prediction_model, conf_level = 0.90, size = 0.2) +
  labs(x = "90% credibility interval")
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults2
#| fig-cap: "Credible intervals for predictors of support for Biden with x_axis limits"
# Create the model plot
model_plot <- modelplot(poverty_prediction_model, conf_level = 0.90, size = 0.2)

# Modify the x-axis limits
model_plot + xlim(-20, 20) +  # Adjust the limits as needed
  labs(x = "90% Credibility Interval")
```

\newpage

# References

### 2024-03-28 scripts/02-data_cleaning.R

@@ -20,7 +20,6 @@ raw_poverty_data <-
   )
 
 # first rename column names to increase readability
-# remove duplicate entries (because all those people live in the same household)
 # filter out ages under 25
 # filter out income is negative
 # poverty status: 1-in poverty; 0-not in poverty
@@ -37,7 +36,6 @@ cleaned_poverty_data <-
     income = spm_totval, 
     age = spm_hage
   ) |>
-  distinct() |>
   filter(age > 25) |>
   filter(income > 0) |>
   mutate(
@@ -53,7 +51,7 @@ cleaned_poverty_data <-
       50000 <= income & income < 99999 ~ "50k-100k",
       100000 <= income & income < 149999 ~ "100k-150k",
       150000 <= income & income < 199999 ~ "150k-200k",
-      200000 <= income & income < 249999 ~ "200k-150k",
+      200000 <= income & income < 249999 ~ "200k-250k",
       250000 <= income ~ "above 250k"
     ),
     age = case_when(

### 2024-03-28 scripts/03-test_data.R

@@ -34,7 +34,11 @@ if (all(test_poverty_data$poverty_status |>
 }
 
 # check the range of data for mortgage
-correct_mortgage <- c("Owner without mortgage", "Owner with mortgage", "Renter")
+correct_mortgage <- 
+  c(
+    "Owner without mortgage", 
+    "Owner with mortgage", 
+    "Renter")
 
 if (all(test_poverty_data$mortgage |>
         unique() %in% correct_mortgage)) 

### 2024-03-29 paper/paper.qmd
adjust graph

@@
 
-posterior_vs_prior(poverty_prediction_model) +
+posterior_vs_prior(poverty_prediction_model, size = 0.3) +
   theme_minimal() +
   theme(legend.position = "bottom") +
+  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
+        text=element_text(size=8)) +
+  guides(color = guide_legend(ncol = 3))+
   coord_flip()

### 2024-03-29 paper/paper.qmd
change to marital status

--- 2024-03-28 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Understanding the Impact of Marital Status, Income, and Age on Individuals' Poverty Status"
+subtitle: "An Analysis of the 2019 Revised Supplemental Poverty Measure Dataset"
 author: 
@@
 
+## Data Source
+
+## Features
+
 ```{r}
+#| include: false
+#| warning: false
+#| message: false
 
+# read in raw data
+raw_poverty_data <-  read_csv(here::here("data/raw_data/raw_poverty_data.csv"))
 ```
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+```{r}
+#| label: tbl-raw
+#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
 
-Talk more about it.
+raw_poverty_data[1:5, 1:4] |>
+  kable()
+```
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+## Data Measurement
 
-Talk way more about it.
+## Methodology
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# read in cleaned data
+poverty2019 <-  read_parquet(here::here("data/analysis_data/cleaned_poverty_data.parquet"))
+
+poverty2019 <-
+  poverty2019 |>
+  mutate(
+    poverty_status = as_factor(poverty_status),
+    marital_status= factor(
+      marital_status,
+      levels = c(
+      "Married - civilian spouse present",
+      "Married - armed forces spouse present",
+      "Married - spouse absent (excluding separated)",
+      "Widowed",
+      "Divorced",
+      "Separated",
+      "Never Married")
+    ),
+    income = factor(
+      income,
+      levels = c(
+      "below 10k",
+      "10k-50k",
+      "50k-100k",
+      "100k-150k",
+      "150k-200k",
+      "200k-250k",
+      "above 250k")
+    ),
+    age = factor(
+      age,
+      levels = c(
+      "25-34",
+      "35-44",
+      "45-54",
+      "55-64",
+      "above 65")
+    )
+  ) |>
+  select(poverty_status, marital_status, income, age)
+```
+
+```{r}
+#| label: tbl-clean-data
+#| tbl-cap: "Preview of the cleaned 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
+
+poverty2019[5:9, 1:4] |>
+  kable()
+```
+
+```{r}
+#| label: tbl-summary
+#| tbl-cap: "Statistics summary of the cleaned 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
+
+poverty2019 |>
+  summary() |>
+  kable(align = "c") %>%
+  kable_styling(font_size = 8)
+```
+
+## Data Visualization
+```{r}
+#| label: fig-mortgage
+#| fig-cap: The distribution of poverty status by marital status
+#| echo: false
+#| warning: false
+#| message: false
+poverty2019 |>
+  ggplot(aes(x = marital_status, fill = poverty_status)) +
+  geom_bar(position = "dodge") +
+  coord_flip() +
+  theme_minimal() +
+  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
+        text=element_text(size=6)) +
+  scale_fill_brewer(palette = "Set1") +
+  labs(
+    x = "Marital status",
+    y = "Number of respondents",
+    fill = "Poverty status"
+  )
+```
+```{r}
+#| label: fig-income
+#| fig-cap: The distribution of poverty status by income level
+#| echo: false
+#| warning: false
+#| message: false
+poverty2019 |>
+  ggplot(aes(x = income, fill = poverty_status)) +
+  geom_bar(position = "dodge") +
+  coord_flip() +
+  theme_minimal() +
+  scale_fill_brewer(palette = "Set1") +
+  labs(
+    x = "Income level",
+    y = "Number of respondents",
+    fill = "Poverty status"
+  )
+```
+
+```{r}
+#| label: fig-age
+#| fig-cap: The distribution of poverty status by age group
+#| echo: false
+#| warning: false
+#| message: false
+poverty2019 |>
+  ggplot(aes(x = age, fill = poverty_status)) +
+  geom_bar(position = "dodge") +
+  coord_flip() +
+  theme_minimal() +
+  scale_fill_brewer(palette = "Set1") +
+  labs(
+    x = "Age group",
+    y = "Number of respondents",
+    fill = "Poverty status"
+  )
+```
 
@@
 # Convert variables to factors
-analysis_data$mortgage <- factor(analysis_data$mortgage)
+analysis_data$marital_status <- factor(analysis_data$marital_status)
 analysis_data$income <- factor(analysis_data$income)
@@
 
-posterior_vs_prior(poverty_prediction_model) +
+posterior_vs_prior(poverty_prediction_model, size = 0.1) +
   theme_minimal() +
   theme(legend.position = "bottom") +
+  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
+        text=element_text(size=5)) +
+  guides(color = guide_legend(ncol = 4))+
   coord_flip()
@@
 #| label: fig-trace1
-#| fig-cap: "Trace plot of intercept and mortgage"
-#| fig-subcap: ["Trace plot of Intercept", "Trace plot of owner without mortgage", "Trace plot of renter"]
+#| fig-cap: "Trace plot of intercept and marital status"
+#| fig-subcap: ["Trace plot of Intercept", "Trace plot of Married - civilian spouse present", "Trace plot of Married - armed forces spouse present", "Trace plot of Married - spouse absent (excluding separated)", "Trace plot of Widowed", "Trace plot of Separated", "Trace plot of Never Married"]
 #| layout-ncol: 2
@@
 plot(poverty_prediction_model, "trace", "(Intercept)")
-plot(poverty_prediction_model, "trace", "mortgageOwner without mortgage")
-plot(poverty_prediction_model, "trace", "mortgageRenter")
+plot(poverty_prediction_model, "trace", "marital_statusMarried - armed forces spouse present")
+plot(poverty_prediction_model, "trace", "marital_statusMarried - civilian spouse present")
+plot(poverty_prediction_model, "trace", "marital_statusMarried - spouse absent (excluding separated)")
+plot(poverty_prediction_model, "trace", "marital_statusNever Married")
+plot(poverty_prediction_model, "trace", "marital_statusSeparated")
+plot(poverty_prediction_model, "trace", "marital_statusWidowed")
 ```
@@
 
-plot(poverty_prediction_model, "rhat")
+plot(poverty_prediction_model, "rhat") + xlim(-100,1.1)
 ```

### 2024-03-29 scripts/01-download_data.R

@@ -18,7 +18,7 @@ raw_poverty_data <- read_dta('https://www2.census.gov/programs-surveys/supplemen
 
 raw_poverty_data <-
   raw_poverty_data |>
-  select(spm_poor, spm_tenmortstatus, spm_totval, spm_hage)
+  select(spm_poor, spm_hmaritalstatus, spm_totval, spm_hage)
 
 
 #### Save data ####

### 2024-03-29 scripts/02-data_cleaning.R

@@ -15,15 +15,13 @@ library(arrow)
 #### Clean data ####
 # read in raw data
 raw_poverty_data <-
-  read_csv(
-    "data/raw_data/raw_poverty_data.csv"
-  )
+  read_csv("data/raw_data/raw_poverty_data.csv")
 
 # first rename column names to increase readability
 # filter out ages under 25
 # filter out income is negative
 # poverty status: 1-in poverty; 0-not in poverty
-# mortgage status: 1-Owner with Mortgage; 2-Owner without Mortgage; 3-Renter
+# marital status: 1 Married - civilian spouse present; 2 Married - armed forces spouse present; 3 Married - spouse absent (excluding separated); 4 Widowed; 5 Divorced; 6 Separated; 7 Never Married
 # mutate income and age variables into categorical variables
   # income level: below 10000, 10000 - 49999, 50000 - 99999, 100000 - 1499999, 150000 - 199999, 200000 - 249999, above 250000
   # age group: 25-34, 35-44, 45-54, 55-64, above 65
@@ -32,7 +30,7 @@ cleaned_poverty_data <-
   raw_poverty_data |>
   rename(
     poverty_status = spm_poor, 
-    mortgage = spm_tenmortstatus, 
+    marital_status = spm_hmaritalstatus, 
     income = spm_totval, 
     age = spm_hage
   ) |>
@@ -40,10 +38,14 @@ cleaned_poverty_data <-
   filter(income > 0) |>
   mutate(
     poverty_status = if_else(poverty_status == 1, "In poverty", "Not in poverty"),
-    mortgage = case_when(
-      mortgage == 1 ~ "Owner with mortgage",
-      mortgage == 2 ~ "Owner without mortgage",
-      mortgage == 3 ~ "Renter",
+    marital_status = case_when(
+      marital_status == 1 ~ "Married - civilian spouse present",
+      marital_status == 2 ~ "Married - armed forces spouse present",
+      marital_status == 3 ~ "Married - spouse absent (excluding separated)",
+      marital_status == 4 ~ "Widowed",
+      marital_status == 5 ~ "Divorced",
+      marital_status == 6 ~ "Separated",
+      marital_status == 7 ~ "Never Married"
     ),
     income = case_when(
       income < 10000 ~ "below 10k",

### 2024-03-29 scripts/03-test_data.R

@@ -15,9 +15,9 @@ library(arrow)
 # read in cleaned data
 test_poverty_data <- read_parquet("data/analysis_data/cleaned_poverty_data.parquet")
 
-# Check column class poverty_status, mortgage, income, age
+# Check column class poverty_status, marital_status, income, age
 class(test_poverty_data$poverty_status) == "character"
-class(test_poverty_data$mortgage) == "character"
+class(test_poverty_data$marital_status) == "character"
 class(test_poverty_data$income) == "character"
 class(test_poverty_data$age) == "character"
 
@@ -33,19 +33,23 @@ if (all(test_poverty_data$poverty_status |>
   "Not all of the poverty_status have been cleaned completely"
 }
 
-# check the range of data for mortgage
-correct_mortgage <- 
+# check the range of data for marital_status
+correct_marital_status <- 
   c(
-    "Owner without mortgage", 
-    "Owner with mortgage", 
-    "Renter")
+    "Married - civilian spouse present",
+    "Married - armed forces spouse present",
+    "Married - spouse absent (excluding separated)",
+    "Widowed",
+    "Divorced",
+    "Separated",
+    "Never Married")
 
-if (all(test_poverty_data$mortgage |>
-        unique() %in% correct_mortgage)) 
+if (all(test_poverty_data$marital_status |>
+        unique() %in% correct_marital_status)) 
 {
-  "The cleaned mortgage match the expected mortgage"
+  "The cleaned marital_status match the expected marital_status"
 } else {
-  "Not all of the mortgage have been cleaned completely"
+  "Not all of the marital_status have been cleaned completely"
 }
 
 # check the range of data for income
@@ -56,7 +60,7 @@ correct_income <-
     "50k-100k",
     "100k-150k",
     "150k-200k",
-    "200k-150k",
+    "200k-250k",
     "above 250k"
   )
 if (all(test_poverty_data$income |>

### 2024-03-29 scripts/04-model.R

@@ -17,7 +17,7 @@ library(testthat)
 analysis_data <- read_parquet("data/analysis_data/cleaned_poverty_data.parquet")
 
 # Convert variables to factors
-analysis_data$mortgage <- factor(analysis_data$mortgage)
+analysis_data$marital_status <- factor(analysis_data$marital_status)
 analysis_data$income <- factor(analysis_data$income)
 analysis_data$age <- factor(analysis_data$age)
 
@@ -33,7 +33,7 @@ poverty_data_reduced <-
 
 poverty_prediction_model <-
   stan_glm(
-    poverty_status ~ mortgage + income + age,
+    poverty_status ~ marital_status + income + age,
     data = poverty_data_reduced,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-03-29 paper/paper.qmd
mortgage and income

--- 2024-03-28 paper/paper.qmd
@@
 
+## Data Source
+
+## Features
+
 ```{r}
+#| include: false
+#| warning: false
+#| message: false
 
+# read in raw data
+raw_poverty_data <-  read_csv(here::here("data/raw_data/raw_poverty_data.csv"))
 ```
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+```{r}
+#| label: tbl-raw
+#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
 
-Talk more about it.
+raw_poverty_data[1:5, 1:4] |>
+  kable()
+```
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+## Data Measurement
 
-Talk way more about it.
+## Methodology
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# read in cleaned data
+poverty2019 <-  read_parquet(here::here("data/analysis_data/cleaned_poverty_data.parquet"))
+
+poverty2019 <-
+  poverty2019 |>
+  mutate(
+    poverty_status = as_factor(poverty_status),
+    mortgage_state= factor(
+      mortgage_state,
+      levels = c(
+      "Owner with Mortgage",
+      "Owner without Mortgage or rent-free",
+      "Renter")
+    ),
+    income = factor(
+      income,
+      levels = c(
+      "below 10k",
+      "10k-50k",
+      "50k-100k",
+      "100k-150k",
+      "150k-200k",
+      "200k-250k",
+      "above 250k")
+    )
+  ) |>
+  select(poverty_status, mortgage_state, income)
+```
+
+```{r}
+#| label: tbl-clean-data
+#| tbl-cap: "Preview of the cleaned 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
+
+poverty2019[5:9, 1:4] |>
+  kable()
+```
+
+```{r}
+#| label: tbl-summary
+#| tbl-cap: "Statistics summary of the cleaned 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
+
+poverty2019 |>
+  summary() |>
+  kable(align = "c") %>%
+  kable_styling(font_size = 8)
+```
+
+## Data Visualization
+```{r}
+#| label: fig-mortgage
+#| fig-cap: The distribution of poverty status by mortgage state
+#| echo: false
+#| warning: false
+#| message: false
+poverty2019 |>
+  ggplot(aes(x = mortgage_state, fill = poverty_status)) +
+  geom_bar(position = "dodge") +
+  coord_flip() +
+  theme_minimal() +
+  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
+        text=element_text(size=6)) +
+  scale_fill_brewer(palette = "Set1") +
+  labs(
+    x = "Mortgage state",
+    y = "Number of respondents",
+    fill = "Poverty status"
+  )
+```
+```{r}
+#| label: fig-income
+#| fig-cap: The distribution of poverty status by income level
+#| echo: false
+#| warning: false
+#| message: false
+poverty2019 |>
+  ggplot(aes(x = income, fill = poverty_status)) +
+  geom_bar(position = "dodge") +
+  coord_flip() +
+  theme_minimal() +
+  scale_fill_brewer(palette = "Set1") +
+  labs(
+    x = "Income level",
+    y = "Number of respondents",
+    fill = "Poverty status"
+  )
+```
 
@@
 This shows that the observed data partially matches with our initial belief and expectation about the poverty status of people in the United States in 2019.
-We can see that for people with income level at "below 10k", the posterior distribution greatly shifts from its prior after we input observed data.
-This is likely suggesting that the observed data for "below 10k" strongly contradicts our initial belief.
-This suggests that the majority of people in US who earns less than a total amount of 10k per household annually is in poverty in 2019.
+We can see that for people with income level at "below 10k" and "10k-50k", the posterior distributions shift from their prior after we input observed data; their distribution not crossing 0 at all.
+This is suggesting that the observed data for "below 10k" and "10k-50k" strongly contradict our initial belief.
+So the majority of people in US who earns less than a total amount of 50k per household annually was in poverty in 2019.
+Also, the intercept (people who are owners with mortgage and earn 100-150k) shifts to the left; its distribution not crossing 0.
+This indicates that the actual observation does not match with our prior belief.
+People in this category was less likely to be in poverty state in 2019 United States.
 
@@
 # Convert variables to factors
-analysis_data$mortgage <- factor(analysis_data$mortgage)
+analysis_data$marital_status <- factor(analysis_data$mortgage_state)
 analysis_data$income <- factor(analysis_data$income)
-analysis_data$age <- factor(analysis_data$age)
 
@@
 
-posterior_vs_prior(poverty_prediction_model) +
+posterior_vs_prior(poverty_prediction_model, size = 0.1) +
   theme_minimal() +
   theme(legend.position = "bottom") +
+  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
+        text=element_text(size=8)) +
+  guides(color = guide_legend(ncol = 4))+
   coord_flip()
@@
 #| label: fig-trace1
-#| fig-cap: "Trace plot of intercept and mortgage"
-#| fig-subcap: ["Trace plot of Intercept", "Trace plot of owner without mortgage", "Trace plot of renter"]
+#| fig-cap: "Trace plot of intercept and marital status"
+#| fig-subcap: ["Trace plot of Intercept", "Trace plot of Owner without Mortgage or rent-free", "Trace plot of Renter"]
 #| layout-ncol: 2
@@
 plot(poverty_prediction_model, "trace", "(Intercept)")
-plot(poverty_prediction_model, "trace", "mortgageOwner without mortgage")
-plot(poverty_prediction_model, "trace", "mortgageRenter")
+plot(poverty_prediction_model, "trace", "mortgage_stateOwner without Mortgage or rent-free")
+plot(poverty_prediction_model, "trace", "mortgage_stateRenter")
 ```
@@
 #| warning: false
-#| label: fig-trace3
-#| fig-cap: "Trace plot of employ"
-#| fig-subcap: ["Trace plot of age 35-44", "Trace plot of age 45-54", "Trace plot of age 55-64", "Trace plot of age above 65"]
-#| layout-ncol: 2
-
-plot(poverty_prediction_model, "trace", "age35-44")
-plot(poverty_prediction_model, "trace", "age45-54")
-plot(poverty_prediction_model, "trace", "age55-64")
-plot(poverty_prediction_model, "trace", "ageabove 65")
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
 #| label: fig-rhat
@@
 #| label: fig-modelresults1
-#| fig-cap: "Credible intervals for predictors of support for Biden"
+#| fig-cap: "Credible intervals for predictors of positive poverty status"
 
@@
 #| label: fig-modelresults2
-#| fig-cap: "Credible intervals for predictors of support for Biden with x_axis limits"
+#| fig-cap: "Credible intervals for predictors of positive poverty status with x_axis limits"
+
 # Create the model plot
@@
 # Modify the x-axis limits
-model_plot + xlim(-20, 20) +  # Adjust the limits as needed
+model_plot + xlim(-5, 5) +  # Adjust the limits as needed
   labs(x = "90% Credibility Interval")

### 2024-03-29 scripts/01-download_data.R

@@ -16,9 +16,10 @@ library(tidyverse)
 #### Download data ####
 raw_poverty_data <- read_dta('https://www2.census.gov/programs-surveys/supplemental-poverty-measure/datasets/2019-revised/spm2019_newmeth.dta')
 
+# Select the feature: h_seq, spm_poor, spm_tenmortstatus, spm_totval
 raw_poverty_data <-
   raw_poverty_data |>
-  select(spm_poor, spm_hmaritalstatus, spm_totval, spm_hage)
+  select(h_seq, spm_poor, spm_tenmortstatus, spm_totval)
 
 
 #### Save data ####

### 2024-03-29 scripts/02-data_cleaning.R

@@ -18,34 +18,27 @@ raw_poverty_data <-
   read_csv("data/raw_data/raw_poverty_data.csv")
 
 # first rename column names to increase readability
-# filter out ages under 25
-# filter out income is negative
+# remove duplicate entries
 # poverty status: 1-in poverty; 0-not in poverty
-# marital status: 1 Married - civilian spouse present; 2 Married - armed forces spouse present; 3 Married - spouse absent (excluding separated); 4 Widowed; 5 Divorced; 6 Separated; 7 Never Married
-# mutate income and age variables into categorical variables
+# mortgage state: 1-Owner with Mortgage; 2-Owner without Mortgage or rent-free; 3-Renter
+# mutate income variable into categorical variable
   # income level: below 10000, 10000 - 49999, 50000 - 99999, 100000 - 1499999, 150000 - 199999, 200000 - 249999, above 250000
-  # age group: 25-34, 35-44, 45-54, 55-64, above 65
+# select features: poverty-status, mortgage_state, income
 
 cleaned_poverty_data <-
   raw_poverty_data |>
   rename(
     poverty_status = spm_poor, 
-    marital_status = spm_hmaritalstatus, 
-    income = spm_totval, 
-    age = spm_hage
+    mortgage_state = spm_tenmortstatus, 
+    income = spm_totval
   ) |>
-  filter(age > 25) |>
-  filter(income > 0) |>
+  unique() |>
   mutate(
     poverty_status = if_else(poverty_status == 1, "In poverty", "Not in poverty"),
-    marital_status = case_when(
-      marital_status == 1 ~ "Married - civilian spouse present",
-      marital_status == 2 ~ "Married - armed forces spouse present",
-      marital_status == 3 ~ "Married - spouse absent (excluding separated)",
-      marital_status == 4 ~ "Widowed",
-      marital_status == 5 ~ "Divorced",
-      marital_status == 6 ~ "Separated",
-      marital_status == 7 ~ "Never Married"
+    mortgage_state = case_when(
+      mortgage_state == 1 ~ "Owner with Mortgage",
+      mortgage_state == 2 ~ "Owner without Mortgage or rent-free",
+      mortgage_state == 3 ~ "Renter"
     ),
     income = case_when(
       income < 10000 ~ "below 10k",
@@ -55,15 +48,9 @@ cleaned_poverty_data <-
[…]

### 2024-03-29 scripts/03-test_data.R

@@ -17,9 +17,8 @@ test_poverty_data <- read_parquet("data/analysis_data/cleaned_poverty_data.parqu
 
 # Check column class poverty_status, marital_status, income, age
 class(test_poverty_data$poverty_status) == "character"
-class(test_poverty_data$marital_status) == "character"
+class(test_poverty_data$mortgage_state) == "character"
 class(test_poverty_data$income) == "character"
-class(test_poverty_data$age) == "character"
 
 # Range of data
 # check the range of data for poverty_status
@@ -33,23 +32,18 @@ if (all(test_poverty_data$poverty_status |>
   "Not all of the poverty_status have been cleaned completely"
 }
 
-# check the range of data for marital_status
-correct_marital_status <- 
+# check the range of data for mortgage_state
+correct_mortgage_state <- 
   c(
-    "Married - civilian spouse present",
-    "Married - armed forces spouse present",
-    "Married - spouse absent (excluding separated)",
-    "Widowed",
-    "Divorced",
-    "Separated",
-    "Never Married")
-
-if (all(test_poverty_data$marital_status |>
-        unique() %in% correct_marital_status)) 
+    "Owner with Mortgage",
+    "Owner without Mortgage or rent-free",
+    "Renter")
+if (all(test_poverty_data$mortgage_state |>
+        unique() %in% correct_mortgage_state)) 
 {
-  "The cleaned marital_status match the expected marital_status"
+  "The cleaned mortgage_state match the expected mortgage_state"
 } else {
-  "Not all of the marital_status have been cleaned completely"
+  "Not all of the mortgage_state have been cleaned completely"
 }
 
 # check the range of data for income
@@ -70,22 +64,3 @@ if (all(test_poverty_data$income |>
 } else {
   "Not all of the income have been cleaned completely"
 }
-
-# check the range of data for age
-correct_age <- 
-  c(
-    "25-34",
-    "35-44",
-    "45-54",
-    "55-64",
-    "above 65"
-  )
-if (all(test_poverty_data$age |>
-        unique() %in% correct_age)) 
-{
-  "The cleaned age match the expected age"
-} else {
-  "Not all of the age have been cleaned completely"
-}
-
-

### 2024-03-29 scripts/04-model.R

@@ -17,9 +17,8 @@ library(testthat)
 analysis_data <- read_parquet("data/analysis_data/cleaned_poverty_data.parquet")
 
 # Convert variables to factors
-analysis_data$marital_status <- factor(analysis_data$marital_status)
+analysis_data$marital_status <- factor(analysis_data$mortgage_state)
 analysis_data$income <- factor(analysis_data$income)
-analysis_data$age <- factor(analysis_data$age)
 
 # Create poverty_status variable in binary form
 analysis_data$poverty_status <- ifelse(analysis_data$poverty_status == "In poverty", 1, 0)
@@ -33,7 +32,7 @@ poverty_data_reduced <-
 
 poverty_prediction_model <-
   stan_glm(
-    poverty_status ~ marital_status + income + age,
+    poverty_status ~ income + mortgage_state,
     data = poverty_data_reduced,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-03-30 paper/paper.qmd
finish model and result

@@
 ---
-title: "Understanding the Impact of Marital Status, Income, and Age on Individuals' Poverty Status"
+title: "Understanding the Impact of Mortgage State and Income on Household's Poverty Status"
 subtitle: "An Analysis of the 2019 Revised Supplemental Poverty Measure Dataset"
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper examines household poverty status in the United States in 2019 using the 2019 Revised Supplemental Poverty Measure (SPM) Research Dataset from the United States Census Bureau.
+Using logistic regression analysis, we investigate the influence of household income level and mortgage state on poverty status.
+Our findings reveal a higher likelihood of poverty for households earning less than $50000 and for renters compared to property owners.
+Conversely, all households with incomes exceeding $100000 are not in poverty.
+These insights emphasize the need for targeted interventions to address socioeconomic inequalities and promote financial well-being among US households."
 format: pdf
+toc: true
 number-sections: true
@@
 
+\newpage
 # Introduction
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+The remainder of this paper is structured as follows. @sec-data....
 
-The remainder of this paper is structured as follows. @sec-data....
+## Estimand
 
@@
 ## Data Source
+This report uses the 2019 Revised Supplemental Poverty Measure (SPM) Research Dataset provided by the United States Census Bureau as our main source of data.
+The Census Bureau provides information about the people and economy of the United States, with a goal to support economic growth, enhance scientific knowledge, and assist in making informed decisions.
+The Supplemental Poverty Measure (SPM), similar to the official poverty measure, tackles the shortcomings of poverty assessment.
+It determines poverty status by gauging resources against a defined standard of living.
+This standard considers expenditures on essentials like food, clothing, shelter, and utilities, plus a bit more for additional expenses.
+The resources assessed include disposable income, accounting for taxes and some noncash benefits, available to cover these needs.
+In 2019, the dataset contained a total of 157,959 entries.
 
 ## Features
+The original SPM 2019 dataset, which shows in @tbl-raw, contains 157959 data entries and many variables.
+Since it is difficult to observe such a large dataset, this report will only explore and analyze through several data features.
+We chose these 4 variables: "h_seq", "spm_poor", "spm_tenmortstatus", "spm_totval", in our analysis.
+
+1. h_seq: Household sequence number, unique identifier for each household.
+
+2. spm_poor: the poverty status; 1 representing "In poverty" and 0 represneting "Not in poverty".
+
+3. spm_tenmortstatus: household's tenure/mortgage status; 1 representing "Owner with Mortgage", 2 representing "Owner without Mortgage or rent-free", and 3 representing "Renter".
+
+4. spm_totval: household's cash income
 
@@
 
+The original dataset includes duplicate responses from the same household.
+As the dataset features are related to the entire household as a whole, this paper will eliminate duplicate entries, using the variable "h_seq" which is an unique household identifier, to ensure data accuracy and consistency.
+
+The dataset is cleaned by renaming of column headers, defining column classes, and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
+After cleaning, 62917 rows of data with 3 data features remain.
+@tbl-clean-data shows a preview of the cleaned dataset.
+
 ```{r}
@@
       "Owner with Mortgage",
-      "Owner without Mortgage or rent-free",
+      "Owner without Mortgage",
       "Renter")
@@
 
-poverty2019[5:9, 1:4] |>
+poverty2019[1:5, 1:3] |>
   kable()
 ```
+
+@tbl-summary is a summary of the cleaned data, showing detailed statistics about the dataset.
+As we can see from the table, there are more households which are not in poverty.
+Respondents cover a wide range of income levels, with "50k-100k" and "100k-150k" being more heavily represented.
+Also, the mortgage states of the households vary, each of the three mortgage states being fairly evenly represented.
 
@@
   kable(align = "c") %>%
-  kable_styling(font_size = 8)
+  kable_styling(font_size = 12)
 ```
@@
 ## Data Visualization
+@fig-mortgage illustrates the relationship between mortgage state and poverty status of households.
+It can be seen that renters constitute the highest proportion of households experiencing poverty, followed by owners without mortgages, and finally owners with mortgages.
+Conversely, for households who are not in poverty, owners with mortgages represent the largest segment, followed by owners without mortgages, and then renters.
+In addition, it can be observed that overall a larger proportion of households are not experiencing poverty.
+Despite variations in poverty rates across different mortgage states, the cumulative data shows economic stability among the majority of households in the United States in 2019.
+
 ```{r}
@@
 poverty2019 |>
-  ggplot(aes(x = mortgage_state, fill = poverty_status)) +
-  geom_bar(position = "dodge") +
-  coord_flip() +
+  ggplot(aes(x = poverty_status, fill = mortgage_state)) +
+  geom_bar(width = 0.2) +
   theme_minimal() +
-  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
-        text=element_text(size=6)) +
   scale_fill_brewer(palette = "Set1") +
   labs(
-    x = "Mortgage state",
+    x = "Poverty Status",
     y = "Number of respondents",
-    fill = "Poverty status"
+    fill = "Mortgage state"
   )
 ```
+
+@fig-income shows the relationship between household annual total income and their poverty status.
+We can see that among households earning more than 100k annually, no household is in poverty.
+For household with income between 50k to 100k, almost all of them are not in poverty with only a very small number of households in poverty.
+Within the income range of 10k-50k, the majority of households are not in poverty, but there are some families in poverty.
+However, for households earning below 10k annually, almost all of them are in poverty.
+This indicates a strong positive correlation between higher income levels and financial stability.
+
 ```{r}
@@
 
+\newpage
 # Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+In our analysis, we utilized a Bayesian logistic regression model to examine the relationship between poverty status and two demographic factors, mortgage status and income of the household.
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
 ## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
+The model is formulated as follows:
 \begin{align} 
 y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
-\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{mortgage}_i + \beta_2 \times \mbox{income}_i + \beta_3 \times \mbox{age}_i\\
+\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{mortgage}_i + \beta_2 \times \mbox{income}_i\\
 \alpha &\sim \mbox{Normal}(0, 2.5) \\
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+In this model, $y_i$ represents the binary outcome variable indicating whether a household is in poverty (as opposed to not in poverty).
+The probability of being in poverty ($\pi_i$) is modeled using a logistic link function (\text{logit}($\pi_i$)), which is a linear combination of the intercept ($\alpha$) and the coefficients ($\beta_1$, $\beta_2$) corresponding to the predictor variables, mortgage state and income level, respectively.
+These predictor variables are denoted as income_i and mortgage_state_i, where $i$ indexes the individual households in the dataset.
+
+The intercept ($\alpha$) and coefficients ($\beta_1$, $\beta_2$) are assigned informative prior distributions to regularize the model.
+Specifically, we assume a normal distribution with a mean of 0 and a standard deviation of 2.5 for each parameter.
+
+We chose this modeling approach for several reasons.
+Firstly, logistic regression is ideal for binary outcome variables, making it appropriate for analyzing poverty status.
+Also, Bayesian methods enable us to integrate prior knowledge and uncertainty into the analysis, resulting in more reliable estimates of the model parameters.
+
+While alternative modeling approaches like linear regression were considered, we selected Bayesian logistic regression due to the binary nature of our outcome variable, poverty status.
+
+We use the `rstanarm` package [@citeRstanarm] in R [@citeR] to run the model.
+Default priors from `rstanarm` is used.
+Rstanarm employs Markov chain Monte Carlo (MCMC) techniques to estimate the posterior distribution of the parameters.
+To avoid excessive runtime, 1000 data entries are randomly sampled to fit the model with random seed 215.
+Model diagnostics, including convergence checks and posterior summaries, are available in the supplementary materials (see Appendix @sec-model-details).
 
 ### Model justification
+Regarding the relationship between mortgage state and poverty state, we anticipate that both owners without mortgages and owners with mortgages are less likely to experience poverty.
+Owners without mortgages typically indicate financial stability, as they have already paid off their homes and can allocate financial resources toward household necessities, other than paying back the mortgage / rent.
+Also, without monthly mortgage or rent payments, their financial burden is reduced, further decreasing the likelihood of poverty.
+Similarly, owners with mortgages demonstrate financial responsibility by securing loans from banks.
+The approval of these loans suggests that the bank is confident in owner's ability to manage repayments, indicating a lower possibility of poverty.
+On the other hand, renters are more likely to experience poverty due to their lack of property ownership and the hidden implication that they do not earn enough to purchase their own property.
+Renters typically have fewer savings and may earn insufficient salaries to cover rental and essential expenses.
+Consequently, the combination of rental payments and other financial obligations increases their vulnerability to poverty.
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+In terms of income levels, we expect a positive relationship between household's income level and their poverty status.
+This expectation arises from common sense that higher income typically reduces the likelihood of experiencing poverty.
+If a family is earning a lot of money, then naturally it would not be in poverty.
+Households that are earning less than 10k annually in the United States obviously have a larger chance in being in poverty due to the high cost of living.
+Their income probably barely covers or cannot cover for essential expenses such as housing, transportation, or food.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+## Model Implication
+For posterior predictive checks, in @fig-post_dist, the great fit of the posterior distribution from our logistic regression model with actual poverty data suggests accurate capture of poverty status patterns.
+This indicates the accuracy of our model's poverty status prediction to the 2019 SPM data.
+Also, @fig-post_prior compares the posterior to the prior, which shows some parameter changes such as "below 10k" and "10k-50k", as well as the intercept (people who are owners with mortgage and earn 100-150k).
+The discrepancy for the parameter suggest that our prior may not be very accurate in regards to these specific aspects.
+
+The trace plots in @fig-trace1 and @fig-trace2 do not suggest anything out of the ordinary.
+Also, with the Rhat plot (@fig-rhat), we can observe that everything is close to 1, and no more than 1.05, which shows the great convergence of Markov chain Monte Carlo for our model.
+
+More detailed explanation of each plot can be found in Appendix @sec-model-details.
 
 # Results
+Our results are summarized in @tbl-modelresults.
+Our results generally matches our expectation.
+To avoid multicollinearity, the model excludes one variable from each category: mortgage state "Owner with mortgage" and income level "100k-150k".
+The intercept represents the estimated log-odds of being in poverty when all other predictors are held constant at their reference levels.
+In this case, the estimated log-odds of being in poverty for people who are owner with mortgage and their total household income level to be 100k-150k annually is −6.782.
 
-Our results are summarized in @tbl-modelresults.
+The possibility of households with income level below 10k being in poverty is large.
+The estimated coefficient of 17.653 suggests that, holding all other variables constant, households with income less than 10k are estimated to have a 17.653 unit increase in the log-odds of being in poverty compared to the reference group.
+Households with income level at "10k-50k" and "50k-100k" on average are also more likely to be in poverty compared to the reference group, with the estimated coefficient to be 5.478 and 1.640 respectively.
+
+The mortgage status of individuals also influences their poverty status.
+As expected, renters exhibit a higher likelihood of experiencing poverty, as shown by the estimated coefficient of 0.359.
+However, it's important to note that while the coefficient is positive compared to the reference group, its magnitude is relatively small, suggesting a moderate rather than a substantial difference.
+
+The mortgage status of individuals also influences their poverty status.
+As expected, renters exhibit a higher likelihood of experiencing poverty, as shown by the estimated coefficient of 0.359.
+However, it's important to note that while the coefficient is positive compared to the reference group, its magnitude is relatively small, suggesting a moderate rather than a substantial difference.
+
+@fig-modelresults1 (see @sec-credibility-interval) shows range of coefficient estimates of our model within the 90% probability.
+Due to the fact that the credibility interval for mortgage "Renter" and "Owners without mortgage" is quite small, it is hard to observe the trend of the 90% credibility intervals of these two variables.
+Therefore we created @fig-modelresults2 with the x axis limited from -5 to 5.
+
+Combining @fig-modelresults1 and @fig-modelresults2, we observe statistical significance for the coefficient estimates for household with income below 10k, household income between 10k to 50k, and the intercept, household which owns a property with mortgage and has a income level between 100k to 150k.
+The estimates are significant because their credibility intervals do not cross 0.
+The value for the estimates are in log-odds, indicating that if the coefficient is positive, the household is in poverty, if negative, the household is not in poverty.
 
@@
 
-## First discussion point {#sec-first-point}
+## Relationship betwen Mortgage State and Poverty Status
 
@@
 
-## Second discussion point
+## Relationship betwen Income and Poverty Status
 
-## Third discussion point
+Average annual expenditure $63,036
+
+## mortgage interest rate etc...
 
@@
 
-@fig-trace1, @fig-trace2, @fig-trace3 are the trace plots of the model.
+@fig-trace1 and @fig-trace2 are the trace plots of the model.
 It tells us if there is existence of signs that the our model runs into issues.
@@
 #| fig-cap: "Trace plot of intercept and marital status"
-#| fig-subcap: ["Trace plot of Intercept", "Trace plot of Owner without Mortgage or rent-free", "Trace plot of Renter"]
+#| fig-subcap: ["Trace plot of Intercept", "Trace plot of Owner without Mortgage", "Trace plot of Renter"]
 #| layout-ncol: 2
@@
 plot(poverty_prediction_model, "trace", "(Intercept)")
-plot(poverty_prediction_model, "trace", "mortgage_stateOwner without Mortgage or rent-free")
+plot(poverty_prediction_model, "trace", "mortgage_stateOwner without Mortgage")
 plot(poverty_prediction_model, "trace", "mortgage_stateRenter")
@@
 ```
+
+## 90% Credibility Interval {#sec-credibility-interval}
 ```{r}
@@
 modelplot(poverty_prediction_model, conf_level = 0.90, size = 0.2) +
-  labs(x = "90% credibility interval")
+  labs(x = "90% credibility interval") +
+  theme(axis.text.y = element_text(size = 6))
 ```
@@
 model_plot + xlim(-5, 5) +  # Adjust the limits as needed
-  labs(x = "90% Credibility Interval")
+  labs(x = "90% Credibility Interval") +
+  theme(axis.text.y = element_text(size = 6))
 ```

### 2024-03-30 scripts/02-data_cleaning.R

@@ -37,7 +37,7 @@ cleaned_poverty_data <-
     poverty_status = if_else(poverty_status == 1, "In poverty", "Not in poverty"),
     mortgage_state = case_when(
       mortgage_state == 1 ~ "Owner with Mortgage",
-      mortgage_state == 2 ~ "Owner without Mortgage or rent-free",
+      mortgage_state == 2 ~ "Owner without Mortgage",
       mortgage_state == 3 ~ "Renter"
     ),
     income = case_when(

### 2024-03-30 scripts/03-test_data.R

@@ -36,7 +36,7 @@ if (all(test_poverty_data$poverty_status |>
 correct_mortgage_state <- 
   c(
     "Owner with Mortgage",
-    "Owner without Mortgage or rent-free",
+    "Owner without Mortgage",
     "Renter")
 if (all(test_poverty_data$mortgage_state |>
         unique() %in% correct_mortgage_state)) 

### 2024-03-30 paper/paper.qmd
abstract

@@
 
-The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is structured into different sections.
+@sec-data demonstrates the data used for our report and includes some tables and graphs to illustrate the different groups of people in our data.
+@sec-model builds the model and discusses its justification and explanation.
+@sec-result highlights the results of the predictions using tables and graphs.
+@sec-discussion contains discussions that conducted based on the findings, which addresses the voting prediction results based on race, region, employment status, and the influence of COVID-19 and in-mail voting systems.
+Statistical programming language R [@citeR] is used in this report, with packages `tidyverse` [@citeTidyverse], `here` [@citeHere], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModleSummary], `ggplot2` [@citeGgplot2], `knitr` [@citeKnitr], `marginaleffects` [@citeMarginalEffects], `plotly` [@citePlotly], `tibble` [@citeTibble], `margins` [@citeMargins], `testthat` [@citetestthat] and `kableExtra` [@citeKableExtra].
 
@@
 
-The dataset is cleaned by renaming of column headers, defining column classes, and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
+The dataset is cleaned by renaming of column headers, defining column classes, grouping the income variable into different income levels (here the assumed facts are that the income unit is in USD and 'k' stands for thousand, eg.
+10k = 10,000), and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
 After cleaning, 62917 rows of data with 3 data features remain.

### 2024-03-30 scripts/00-simulate_data.R
simulate data

@@ -1,19 +1,43 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Simulates the US 2019 poverty status data
+# Author: Mary Cheng
+# Date: 29 March 2024
+# Contact: maryc.cheng@mail.utoronto.ca
+# License: --
+# Pre-requisites: --
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+# set seed
+set.seed(215)
 
+num_obs <- 200
 
-
+# simulate poverty status data based on mortgage state and income level
+simulate_poverty_status <- 
+  tibble(
+    mortgage_state = sample(0:2, num_obs, replace = TRUE, prob = c(0.33, 0.33, 0.34)), # Simulate mortgage state (0 = without mortgage, 1 = with mortgage, 2 = renter)
+    income = rnorm(num_obs, mean = 60000, sd = 20000), # Simulate income level
+    poverty_status = ifelse(
+      (mortgage_state == 0 & income < 40000) | 
+        (mortgage_state == 1 & income < 30000) | 
+        (mortgage_state == 2 & income < 25000), 1, 0) # Simulate poverty status based on mortgage state and income level
+  ) |>
+  mutate(
+    poverty_status = if_else(poverty_status == 1, "In poverty", "Not in poverty"),
+    mortgage_state = case_when(
+      mortgage_state == 0 ~ "Owner without mortgage",
+      mortgage_state == 1 ~ "Owner with mortgage",
+      mortgage_state == 2 ~ "Renter"
+    ),
+    income = case_when(
+      income < 50000 ~ "below 50k",
+      50000 <= income & income < 60000 ~ "50k-60k",
+      60000 <= income & income < 70000 ~ "60k-70k",
+      70000 <= income ~ "above 70k"
+    )) |>
+  select(poverty_status, mortgage_state, income)
+  

### 2024-03-30 paper/paper.qmd
some intro done

@@
 Using logistic regression analysis, we investigate the influence of household income level and mortgage state on poverty status.
-Our findings reveal a higher likelihood of poverty for households earning less than $50000 and for renters compared to property owners.
-Conversely, all households with incomes exceeding $100000 are not in poverty.
+Our findings reveal a higher likelihood of poverty for households' total annual earning less than $50000 and for renters compared to property owners.
+All households with incomes exceeding $100000 are not in poverty.
 These insights emphasize the need for targeted interventions to address socioeconomic inequalities and promote financial well-being among US households."
@@
 
+Therefore, it is important for us to analyze why some US household are in poverty in 2019 and how various demographic and financial factors may has an influence on it.
+Perhaps then we could understand how to reduce the number of households in poverty.
+We analyze the 2019 Revised SPM Research dataset from the United States Census Bureau (USCB) [@citeCes2020].
+USCB is a principal agency of the U.S.
+Federal Statistical System, responsible for producing data about the American people and economy.
+We build a prediction model for households poverty status using 2019 Revised SPM dataset.
+This allows us to discover trends and address potential economic and societal situations that may cause the poverty status for some US households in 2019.
+
+In this paper, a logistic regression model is used to predict the poverty status of households in the United States in 2019, with data from the 2019 Revised Supplemental Poverty Measure Research dataset.
+Logistic regression is a great choice since it is used to predict binary outcomes, such as poverty status (in poverty or not in poverty).
+Our analysis focuses on estimating the likelihood of household being in poverty, based on various demographic and financial factors captured in the SPM dataset.
+We selected data features: mortgage state and household annual total income.
+The estimand in this paper is the number of households who are in poverty in reality.
+However, it is difficult to measure the exact number of households who are in poverty since there are millions of people in the United States and not all of them will be assessed due to various difficulties.
+Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCB.
+
+The logistic regression model shows that in the US in 2019, households with an annual total income less than $50000 are almost all in poverty.
+For households earning $50000 to $100000, the majority is not in poverty, yet a notable amount of them are still in poverty.
+For households earning above $100000, all of them are not in poverty.
+Additionally, renters are more likely to be in poverty compared to owners with or without mortgage.
+
 The remainder of this paper is structured into different sections.
@@
 @sec-result highlights the results of the predictions using tables and graphs.
-@sec-discussion contains discussions that conducted based on the findings, which addresses the voting prediction results based on race, region, employment status, and the influence of COVID-19 and in-mail voting systems.
+@sec-discussion contains discussions that conducted based on the findings, which addresses the poverty status results based on mortgage states and income levels.
 Statistical programming language R [@citeR] is used in this report, with packages `tidyverse` [@citeTidyverse], `here` [@citeHere], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModleSummary], `ggplot2` [@citeGgplot2], `knitr` [@citeKnitr], `marginaleffects` [@citeMarginalEffects], `plotly` [@citePlotly], `tibble` [@citeTibble], `margins` [@citeMargins], `testthat` [@citetestthat] and `kableExtra` [@citeKableExtra].
@@
 ## Estimand
+The estimand in this paper is the number of households who are in poverty in reality.
+However, it is difficult to measure the exact number of households who are in poverty since there are millions of people in the United States and not all of them will be assessed due to various difficulties.
+Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCS.
 
@@
     x = "Poverty Status",
-    y = "Number of respondents",
+    y = "Number of households",
     fill = "Mortgage state"
@@
     x = "Income level",
-    y = "Number of respondents",
+    y = "Number of households",
     fill = "Poverty status"

### 2024-03-31 paper/paper.qmd
add ref

@@
 ---
-title: "Understanding the Impact of Mortgage State and Income on Household's Poverty Status"
-subtitle: "An Analysis of the 2019 Revised Supplemental Poverty Measure Dataset"
+title: "Understanding the Impact of Mortgage State and Income on Household Poverty Status"
+subtitle: "An Analysis of Household Poverty Status in the United States in 2019"
 author: 
@@
 # Introduction
-
+Over the last few decades, addressing poverty has remained a persistent issue in the United States, leading to the implementation of multiple policies in an effort to mitigate this problem [@Kearney_Harris].
+In 2019, the United States experienced the culmination of an economic expansion that followed the Great Recession, marking the last full year before the onset of COVID-19 pandemic.
+The robustness of the U.S.
+economy during this period was evidenced by low poverty rates, indicating a notable decrease in economic hardship characterized by low income.
+The poverty rate in the United States was estimated to be 10.5% in 2019, down from 11.8% in 2018, marking the fifth consecutive year of decline [@Semega_2022].
+Despite various interventions and economic policies, understanding the intricate dynamics influencing household poverty status is still difficult.
 Therefore, it is important for us to analyze why some US household are in poverty in 2019 and how various demographic and financial factors may has an influence on it.
 Perhaps then we could understand how to reduce the number of households in poverty.
-We analyze the 2019 Revised SPM Research dataset from the United States Census Bureau (USCB) [@citeCes2020].
-USCB is a principal agency of the U.S.
-Federal Statistical System, responsible for producing data about the American people and economy.
+We analyze the 2019 Revised SPM Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
+USCB is an agency of the U.S.
+Federal Statistical System, responsible for producing data about the American people and economy [@uscb].
 We build a prediction model for households poverty status using 2019 Revised SPM dataset.
@@
 @fig-mortgage illustrates the relationship between mortgage state and poverty status of households.
-It can be seen that renters constitute the highest proportion of households experiencing poverty, followed by owners without mortgages, and finally owners with mortgages.
+It can be seen that renters are the highest proportion of households experiencing poverty, followed by owners without mortgages, and finally owners with mortgages.
 Conversely, for households who are not in poverty, owners with mortgages represent the largest segment, followed by owners without mortgages, and then renters.
@@
 poverty2019 |>
-  ggplot(aes(x = poverty_status, fill = mortgage_state)) +
-  geom_bar(width = 0.2) +
+  ggplot(aes(x = mortgage_state, fill = poverty_status)) +
+  geom_bar(position = "dodge", width = 0.5) +
   theme_minimal() +
+  coord_flip() +
   scale_fill_brewer(palette = "Set1") +
   labs(
-    x = "Poverty Status",
+    x = "Mortgage state",
     y = "Number of households",
-    fill = "Mortgage state"
+    fill = "Poverty Status"
   )
 ```
+
+#```{r}
+##| label: fig-mortgage
+##| fig-cap: The distribution of poverty status by mortgage state
+##| echo: false
+##| warning: false
+##| message: false
+#poverty2019 |>
+#  ggplot(aes(x = poverty_status, fill = mortgage_state)) +
+#  geom_bar(width = 0.2) +
+#  theme_minimal() +
+#  scale_fill_brewer(palette = "Set1") +
+#  labs(
+#    x = "Poverty Status",
+#    y = "Number of households",
+#    fill = "Mortgage state"
+#  )
+#```
 
@@
   ggplot(aes(x = income, fill = poverty_status)) +
-  geom_bar(position = "dodge") +
+  geom_bar(position = "dodge", width = 0.5) +
   coord_flip() +
@@
 \newpage
-# Model
+
+# Model {#sec-model}
 In our analysis, we utilized a Bayesian logistic regression model to examine the relationship between poverty status and two demographic factors, mortgage status and income of the household.
@@
 
-# Results
+# Results {#sec-result}
 Our results are summarized in @tbl-modelresults.
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2024-03-31 paper/paper.qmd
check point

@@
 ## Data Source
-This report uses the 2019 Revised Supplemental Poverty Measure (SPM) Research Dataset provided by the United States Census Bureau as our main source of data.
+This report uses the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset provided by the United States Census Bureau as our main source of data.
 The Census Bureau provides information about the people and economy of the United States, with a goal to support economic growth, enhance scientific knowledge, and assist in making informed decisions.
@@
 This standard considers expenditures on essentials like food, clothing, shelter, and utilities, plus a bit more for additional expenses.
-The resources assessed include disposable income, accounting for taxes and some noncash benefits, available to cover these needs.
+The resources assessed include disposable income, accounting for taxes and some non-cash benefits, available to cover these needs.
 In 2019, the dataset contained a total of 157,959 entries.

### 2024-03-31 paper/paper.qmd
add some features

@@
 Perhaps then we could understand how to reduce the number of households in poverty.
-We analyze the 2019 Revised SPM Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
+We analyze the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
 USCB is an agency of the U.S.
@@
 We build a prediction model for households poverty status using 2019 Revised SPM dataset.
+Supplemental Poverty Measure was developed due to the continued criticism towards the official poverty measure in the US.
+SPM addresses numerous concerns of official measure critics, and its intent is to provide an improved statistical picture of poverty [@Gesumaria_2015].
 This allows us to discover trends and address potential economic and societal situations that may cause the poverty status for some US households in 2019.
@@
 The Census Bureau provides information about the people and economy of the United States, with a goal to support economic growth, enhance scientific knowledge, and assist in making informed decisions.
+The official poverty measure has been criticized due to its multiple limitations.
+For example, it does not consider benefits from most of the largest programs that aid the low-income population.
+For instance, it uses money income before taxes, so it does not measure the income available for individuals to spend, which for most people is after-tax income.
+Therefore, any effects of tax credits designed to assist persons with low income are not considered by the official measure.
 The Supplemental Poverty Measure (SPM), similar to the official poverty measure, tackles the shortcomings of poverty assessment.
@@
 This standard considers expenditures on essentials like food, clothing, shelter, and utilities, plus a bit more for additional expenses.
-The resources assessed include disposable income, accounting for taxes and some non-cash benefits, available to cover these needs.
+The resources assessed include disposable income, accounting for taxes and some non-cash benefits, available to cover these needs [@Gesumaria_2015].
 In 2019, the dataset contained a total of 157,959 entries.
@@
 Since it is difficult to observe such a large dataset, this report will only explore and analyze through several data features.
-We chose these 4 variables: "h_seq", "spm_poor", "spm_tenmortstatus", "spm_totval", in our analysis.
+We first chose these 17 variables: h_seq, spm_poor, spm_tenmortstatus, spm_totval, spm_snapsub, spm_caphousesub, spm_schlunch, spm_engval, spm_wicval, spm_fedtax, spm_eitc, spm_actc, spm_fica, spm_sttax, spm_childsuppd, spm_capwkccxpns, and spm_medxpns.
+This is because we want to calculate the income after tax, plus all the subsidies and benefits, minus all the expenses, just like how the SPM does it.
+If we only choose income before tax to analyze, then our analysis will be like the official poverty measure, which requires improvement.
 
@@
 
-4. spm_totval: household's cash income
+4. spm_totval: household's cash income, unit in USD.
+
+5. spm_snapsub: Supplemental Nutrition Assistance Program (SNAP) subsidy, unit in USD.
+
+6. spm_caphousesub: capped housing subsidy, unit in USD.
+
+7. spm_schlunch: school lunch subsidy, unit in USD.
+
+8. spm_engval: energy subsidy, unit in USD.
+
+9. spm_wicval: Women, Infants, and Children subsidy, unit in USD.
+
+10. spm_fedtax: Federal tax, unit in USD.
+
+11. spm_eitc: Federal Earned Income Tax Credit, unit in USD.
+
+12. spm_actc: Additional Child Tax Credit, unit in USD.
+
+13. spm_fica: Federal Insurance Contributions Act and federal retirement contribution, unit in USD.
+
+14. spm_sttax: state tax, unit in USD.
+
+15. spm_childsuppd: child support paid, unit in USD.
+
+16. spm_capwkccxpns: capped work and child care expense, unit in USD.
+
+17. spm_medxpns: Medical Out-Of-Pocket (MOOP) and Medicare Part B subsidy, unit in USD.
 
@@
 
-raw_poverty_data[1:5, 1:4] |>
+raw_poverty_data[1:5, 1:6] |>
+  kable()
+raw_poverty_data[1:5, 7:13] |>
+  kable()
+raw_poverty_data[1:5, 13:17] |>
   kable()
@@
 
-The dataset is cleaned by renaming of column headers, defining column classes, grouping the income variable into different income levels (here the assumed facts are that the income unit is in USD and 'k' stands for thousand, eg.
-10k = 10,000), and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
-After cleaning, 62917 rows of data with 3 data features remain.
+First, a new column "income after tax" is created by summing up all incomes, credits, and benefits and subtracting all expenses.
+Then the dataset is cleaned by renaming of column headers, defining column classes, grouping the income variable into different income levels (here the assumed facts are that the income unit is in USD and 'k' stands for thousand, eg.
+5k = 10,000), and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
+After cleaning, 62990 rows of data with 3 data features remain.
 @tbl-clean-data shows a preview of the cleaned dataset.
@@
       levels = c(
-      "below 10k",
-      "10k-50k",
+      "below 5k",
+      "5k-50k",
       "50k-100k",
@@
 
-#```{r}
-##| label: fig-mortgage
-##| fig-cap: The distribution of poverty status by mortgage state
-##| echo: false
-##| warning: false
-##| message: false
-#poverty2019 |>
-#  ggplot(aes(x = poverty_status, fill = mortgage_state)) +
-#  geom_bar(width = 0.2) +
-#  theme_minimal() +
-#  scale_fill_brewer(palette = "Set1") +
-#  labs(
-#    x = "Poverty Status",
-#    y = "Number of households",
-#    fill = "Mortgage state"
-#  )
-#```
-
 @fig-income shows the relationship between household annual total income and their poverty status.
@@
 For household with income between 50k to 100k, almost all of them are not in poverty with only a very small number of households in poverty.
-Within the income range of 10k-50k, the majority of households are not in poverty, but there are some families in poverty.
-However, for households earning below 10k annually, almost all of them are in poverty.
+Within the income range of 5k-50k, the majority of households are not in poverty, but there are some families in poverty.
+However, for households earning below 5k annually, almost all of them are in poverty.
 This indicates a strong positive correlation between higher income levels and financial stability.
@@
 If a family is earning a lot of money, then naturally it would not be in poverty.
-Households that are earning less than 10k annually in the United States obviously have a larger chance in being in poverty due to the high cost of living.
+Households that are earning less than 5k annually in the United States obviously have a larger chance in being in poverty due to the high cost of living.
 Their income probably barely covers or cannot cover for essential expenses such as housing, transportation, or food.
@@
 This indicates the accuracy of our model's poverty status prediction to the 2019 SPM data.
-Also, @fig-post_prior compares the posterior to the prior, which shows some parameter changes such as "below 10k" and "10k-50k", as well as the intercept (people who are owners with mortgage and earn 100-150k).
+Also, @fig-post_prior compares the posterior to the prior, which shows some parameter changes such as "below 5k" and "5k-50k", as well as the intercept (people who are owners with mortgage and earn 100-150k).
 The discrepancy for the parameter suggest that our prior may not be very accurate in regards to these specific aspects.
@@
 The intercept represents the estimated log-odds of being in poverty when all other predictors are held constant at their reference levels.
-In this case, the estimated log-odds of being in poverty for people who are owner with mortgage and their total household income level to be 100k-150k annually is −6.782.
+In this case, the estimated log-odds of being in poverty for people who are owner with mortgage and their total household income level to be 100k-150k annually is −6.896.
 
-The possibility of households with income level below 10k being in poverty is large.
-The estimated coefficient of 17.653 suggests that, holding all other variables constant, households with income less than 10k are estimated to have a 17.653 unit increase in the log-odds of being in poverty compared to the reference group.
-Households with income level at "10k-50k" and "50k-100k" on average are also more likely to be in poverty compared to the reference group, with the estimated coefficient to be 5.478 and 1.640 respectively.
+The possibility of households with income level below 5k being in poverty is large.
+The estimated coefficient of 17.352 suggests that, holding all other variables constant, households with income less than 5k are estimated to have a 17.352 unit increase in the log-odds of being in poverty compared to the reference group.
+Households with income level at "5k-50k" and "50k-100k" on average are also more likely to be in poverty compared to the reference group, with the estimated coefficient to be 5.346 and 1.502 respectively.
 
 The mortgage status of individuals also influences their poverty status.
-As expected, renters exhibit a higher likelihood of experiencing poverty, as shown by the estimated coefficient of 0.359.
-However, it's important to note that while the coefficient is positive compared to the reference group, its magnitude is relatively small, suggesting a moderate rather than a substantial difference.
-
-The mortgage status of individuals also influences their poverty status.
-As expected, renters exhibit a higher likelihood of experiencing poverty, as shown by the estimated coefficient of 0.359.
+As expected, renters exhibit a higher likelihood of experiencing poverty, as shown by the estimated coefficient of 0.645.
 However, it's important to note that while the coefficient is positive compared to the reference group, its magnitude is relatively small, suggesting a moderate rather than a substantial difference.
@@
 
-Combining @fig-modelresults1 and @fig-modelresults2, we observe statistical significance for the coefficient estimates for household with income below 10k, household income between 10k to 50k, and the intercept, household which owns a property with mortgage and has a income level between 100k to 150k.
+Combining @fig-modelresults1 and @fig-modelresults2, we observe statistical significance for the coefficient estimates for household with income below 5k, household income between 5k to 50k, renters, and the intercept, household which owns a property with mortgage and has a income level between 100k to 150k.
 The estimates are significant because their credibility intervals do not cross 0.
@@
 This shows that the observed data partially matches with our initial belief and expectation about the poverty status of people in the United States in 2019.
-We can see that for people with income level at "below 10k" and "10k-50k", the posterior distributions shift from their prior after we input observed data; their distribution not crossing 0 at all.
-This is suggesting that the observed data for "below 10k" and "10k-50k" strongly contradict our initial belief.
+We can see that for people with income level at "below 5k" and "5k-50k" and "Renters", the posterior distributions shift from their prior after we input observed data; their distribution not crossing 0 at all.
+This is suggesting that the observed data for "below 5k" and "5k-50k" strongly contradict our initial belief.
 So the majority of people in US who earns less than a total amount of 50k per household annually was in poverty in 2019.
@@
 #| fig-cap: "Trace plot of income"
-#| fig-subcap: ["Trace plot of income 10k-50k", "Trace plot of income 150k-200k", "Trace plot of income 200k-250k", "Trace plot of income 50k-100k", "Trace plot of income above 250k", "Trace plot of income below 10k"]
+#| fig-subcap: ["Trace plot of income 5k-50k", "Trace plot of income 150k-200k", "Trace plot of income 200k-250k", "Trace plot of income 50k-100k", "Trace plot of income above 250k", "Trace plot of income below 5k"]
 #| layout-ncol: 2
 
-plot(poverty_prediction_model, "trace", "income10k-50k")
+plot(poverty_prediction_model, "trace", "income5k-50k")
 plot(poverty_prediction_model, "trace", "income150k-200k")
@@
 plot(poverty_prediction_model, "trace", "incomeabove 250k")
-plot(poverty_prediction_model, "trace", "incomebelow 10k")
+plot(poverty_prediction_model, "trace", "incomebelow 5k")
 ```

### 2024-03-31 scripts/01-download_data.R

@@ -19,7 +19,7 @@ raw_poverty_data <- read_dta('https://www2.census.gov/programs-surveys/supplemen
 # Select the feature: h_seq, spm_poor, spm_tenmortstatus, spm_totval
 raw_poverty_data <-
   raw_poverty_data |>
-  select(h_seq, spm_poor, spm_tenmortstatus, spm_totval)
+  select(h_seq, spm_poor, spm_tenmortstatus, spm_totval, spm_snapsub, spm_caphousesub, spm_schlunch, spm_engval, spm_wicval, spm_fedtax, spm_eitc, spm_actc, spm_fica, spm_sttax, spm_childsuppd, spm_capwkccxpns, spm_medxpns)
 
 
 #### Save data ####

### 2024-03-31 scripts/02-data_cleaning.R

@@ -17,20 +17,23 @@ library(arrow)
 raw_poverty_data <-
   read_csv("data/raw_data/raw_poverty_data.csv")
 
+# Calculate money income after tax, minus expenses, plus credits and benefits
+raw_poverty_data$income_after_tax <- raw_poverty_data$spm_totval + raw_poverty_data$spm_snapsub + raw_poverty_data$spm_caphousesub + raw_poverty_data$spm_schlunch + raw_poverty_data$spm_engval + raw_poverty_data$spm_wicval - raw_poverty_data$spm_fedtax + raw_poverty_data$spm_eitc + raw_poverty_data$spm_actc - raw_poverty_data$spm_fica - raw_poverty_data$spm_sttax - raw_poverty_data$spm_childsuppd - raw_poverty_data$spm_capwkccxpns - raw_poverty_data$spm_medxpns
+
 # first rename column names to increase readability
 # remove duplicate entries
 # poverty status: 1-in poverty; 0-not in poverty
 # mortgage state: 1-Owner with Mortgage; 2-Owner without Mortgage or rent-free; 3-Renter
 # mutate income variable into categorical variable
-  # income level: below 10000, 10000 - 49999, 50000 - 99999, 100000 - 1499999, 150000 - 199999, 200000 - 249999, above 250000
-# select features: poverty-status, mortgage_state, income
+  # income level: below 5000, 5000 - 49999, 50000 - 99999, 100000 - 1499999, 150000 - 199999, 200000 - 249999, above 250000
+# remove any NA, select features: poverty-status, mortgage_state, income
 
 cleaned_poverty_data <-
   raw_poverty_data |>
   rename(
     poverty_status = spm_poor, 
     mortgage_state = spm_tenmortstatus, 
-    income = spm_totval
+    income = income_after_tax
   ) |>
   unique() |>
   mutate(
@@ -41,8 +44,8 @@ cleaned_poverty_data <-
       mortgage_state == 3 ~ "Renter"
     ),
     income = case_when(
-      income < 10000 ~ "below 10k",
-      10000 <= income & income < 49999 ~ "10k-50k",
+      income < 5000 ~ "below 5k",
+      5000 <= income & income < 49999 ~ "5k-50k",
       50000 <= income & income < 99999 ~ "50k-100k",
       100000 <= income & income < 149999 ~ "100k-150k",
       150000 <= income & income < 199999 ~ "150k-200k",
[…]

### 2024-03-31 scripts/03-test_data.R

@@ -49,8 +49,8 @@ if (all(test_poverty_data$mortgage_state |>
 # check the range of data for income
 correct_income <- 
   c(
-    "below 10k",
-    "10k-50k",
+    "below 5k",
+    "5k-50k",
     "50k-100k",
     "100k-150k",
     "150k-200k",

### 2024-03-31 scripts/04-model.R

@@ -32,7 +32,7 @@ poverty_data_reduced <-
 
 poverty_prediction_model <-
   stan_glm(
-    poverty_status ~ income + mortgage_state,
+    poverty_status ~ mortgage_state+income,
     data = poverty_data_reduced,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-04-01 paper/paper.qmd
add discussion

@@
 abstract: "This paper examines household poverty status in the United States in 2019 using the 2019 Revised Supplemental Poverty Measure (SPM) Research Dataset from the United States Census Bureau.
-Using logistic regression analysis, we investigate the influence of household income level and mortgage state on poverty status.
+Using logistic regression analysis, we investigate the influence of household income level after tax and mortgage state on poverty status.
 Our findings reveal a higher likelihood of poverty for households' total annual earning less than $50000 and for renters compared to property owners.
@@
 # Introduction
-Over the last few decades, addressing poverty has remained a persistent issue in the United States, leading to the implementation of multiple policies in an effort to mitigate this problem [@Kearney_Harris].
-In 2019, the United States experienced the culmination of an economic expansion that followed the Great Recession, marking the last full year before the onset of COVID-19 pandemic.
+Over the last few decades, addressing poverty has remained a persistent issue in the United States, leading to the implementation of many policies in an effort to mitigate this problem [@Kearney_Harris].
+In 2019, the United States experienced the culmination of an economic expansion that followed the Great Recession.
 The robustness of the U.S.
-economy during this period was evidenced by low poverty rates, indicating a notable decrease in economic hardship characterized by low income.
+economy during this period was evidenced by low poverty rates, indicating a notable decrease in economic hardship.
 The poverty rate in the United States was estimated to be 10.5% in 2019, down from 11.8% in 2018, marking the fifth consecutive year of decline [@Semega_2022].
-Despite various interventions and economic policies, understanding the intricate dynamics influencing household poverty status is still difficult.
-Therefore, it is important for us to analyze why some US household are in poverty in 2019 and how various demographic and financial factors may has an influence on it.
-Perhaps then we could understand how to reduce the number of households in poverty.
+In order to reduce poverty further, it is important for us to analyze why some US household are in poverty in 2019 and how various demographic and financial factors may has an influence on it.
 We analyze the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
@@
 Federal Statistical System, responsible for producing data about the American people and economy [@uscb].
+Supplemental Poverty Measure was developed to address numerous concerns of official poverty measure critics, and its intent is to provide an improved statistical picture of poverty [@Gesumaria_2015].
 We build a prediction model for households poverty status using 2019 Revised SPM dataset.
-Supplemental Poverty Measure was developed due to the continued criticism towards the official poverty measure in the US.
-SPM addresses numerous concerns of official measure critics, and its intent is to provide an improved statistical picture of poverty [@Gesumaria_2015].
-This allows us to discover trends and address potential economic and societal situations that may cause the poverty status for some US households in 2019.
+This allows us to discover trends and address potential financial and societal situations that may cause the poverty status for some US households in 2019.
 
@@
 Our analysis focuses on estimating the likelihood of household being in poverty, based on various demographic and financial factors captured in the SPM dataset.
-We selected data features: mortgage state and household annual total income.
+We select 17 data features for analysis.
+After some data cleaning, we summarize 17 features into 2 variables: 1) mortgage state and 2) household annual total income after tax, plus benefits and minus expenses.
 The estimand in this paper is the number of households who are in poverty in reality.
 However, it is difficult to measure the exact number of households who are in poverty since there are millions of people in the United States and not all of them will be assessed due to various difficulties.
-Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCB.
+Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCS.
 
@@
 
-## Estimand
-The estimand in this paper is the number of households who are in poverty in reality.
-However, it is difficult to measure the exact number of households who are in poverty since there are millions of people in the United States and not all of them will be assessed due to various difficulties.
-Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCS.
-
 # Data {#sec-data}
@@
 The Census Bureau provides information about the people and economy of the United States, with a goal to support economic growth, enhance scientific knowledge, and assist in making informed decisions.
-The official poverty measure has been criticized due to its multiple limitations.
+The official poverty measure has been criticized due to its multiple limitations [@sgp2019].
 For example, it does not consider benefits from most of the largest programs that aid the low-income population.
@@
 We first chose these 17 variables: h_seq, spm_poor, spm_tenmortstatus, spm_totval, spm_snapsub, spm_caphousesub, spm_schlunch, spm_engval, spm_wicval, spm_fedtax, spm_eitc, spm_actc, spm_fica, spm_sttax, spm_childsuppd, spm_capwkccxpns, and spm_medxpns.
-This is because we want to calculate the income after tax, plus all the subsidies and benefits, minus all the expenses, just like how the SPM does it.
+This is because we want to calculate the income after tax, plus all the subsidies and benefits, minus all the expenses, just like how the SPM does it [@sgp2019].
 If we only choose income before tax to analyze, then our analysis will be like the official poverty measure, which requires improvement.
@@
 As we can see from the table, there are more households which are not in poverty.
-Respondents cover a wide range of income levels, with "50k-100k" and "100k-150k" being more heavily represented.
+Respondents cover a wide range of income levels, with "5k-50k" and "50k-100k" being more heavily represented.
 Also, the mortgage states of the households vary, each of the three mortgage states being fairly evenly represented.
@@
 Within the income range of 5k-50k, the majority of households are not in poverty, but there are some families in poverty.
-However, for households earning below 5k annually, almost all of them are in poverty.
+However, for households earning below 5k annually, all of them are in poverty.
 This indicates a strong positive correlation between higher income levels and financial stability.
@@
 #| message: false
+
 poverty2019 |>
   ggplot(aes(x = income, fill = poverty_status)) +
-  geom_bar(position = "dodge", width = 0.5) +
+  geom_bar(position = position_dodge2(width = 0.5, preserve = "single")) +
   coord_flip() +
@@
 ```
+\newpage
 
@@
 ## Relationship betwen Mortgage State and Poverty Status
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+In the USA in 2019, housing costs accounted for more than 30% of household expenditure, making mortgage status a critical factor in determining a family's risk of falling into poverty [@Statistics_2020].
+With 11.8% of the total population (38.3 million people) living below the poverty line, it's evident that housing affordability significantly impacts economic stability.
+The percentage of individuals below the poverty threshold, adjusted for housing costs are: $29,080 for a family of four with a mortgage, $24,413 for a family of four without a mortgage, and $29,194 for a family of four [@Khattar_Pathak_Schweitzer_Khan_McManus_Jones_Benson_Kelly_2023].
+From this, we can assumed that families without mortgages experienced the lowest housing costs, leaving more funds available for other essentials and eliminating the need for alternative housing options.
+In contrast, families with mortgages may face burdensome monthly payments, especially if they have chosen a floating interest rate, making them vulnerable to fluctuations in interest rates.
+Renters also encounter challenges, as rent increases can strain finances and necessitate either accepting higher housing costs or undertaking the time-consuming and costly process of finding new accommodations.
+These housing-related challenges exacerbate the risk of poverty for affected families, underscoring the importance of addressing housing affordability as a key component of poverty alleviation efforts.
 
 ## Relationship betwen Income and Poverty Status
+The relationship between a household's income after tax and its poverty status is characterized by a negative correlation, wherein higher income corresponds to a reduced probability of experiencing poverty.
+This correlation shows the role of financial resources in mitigating economic hardship and facilitating socio-economic well-being.
+However, in the United States, the cost of living is notably high, low-income households often struggle to meet basic needs, resulting in an unending cycle of poverty.
+In 2019, the poverty guideline for a one-person household was set at $12,490 annually [@ASPE].
+So, it is evident that families earning less than $5,000 are definitely living below the poverty line.
+Conversely, individuals with high incomes are less likely to experience poverty, as their financial resources enable them to comfortably cover living expenses.
+They will be able to have savings accumulation, further consolidating their financial security and contributing to wealth accumulation over time.
 
-Average annual expenditure $63,036
+For household in poverty, escaping poverty is tough, but the US government and communities are stepping up to help.
+For example, single parents often struggle with low incomes.
+But they can get help from government programs such as the Supplemental Nutrition Assistance Program (SNAP) [@snap], commonly known as food stamps, which provides eligible low-income individuals and families with funds to purchase nutritious food.
+They can also receive child support payments to help cover expenses for their kids.
+Families with low incomes and lots of kids can also get help.
+Programs like the Child Tax Credit (CTC) [@ctc], which provides tax credits to eligible families with qualifying children.
+Other programs, like housing assistance and Medicaid, provide support with things like affordable housing and healthcare.
+These programs are all aimed at helping families in need to get back on their feet and improve their poverty situation.
 

### 2024-04-02 paper/paper.qmd
fix data feature

@@
 
-1. h_seq: Household sequence number, unique identifier for each household.
-
-2. spm_poor: the poverty status; 1 representing "In poverty" and 0 represneting "Not in poverty".
-
-3. spm_tenmortstatus: household's tenure/mortgage status; 1 representing "Owner with Mortgage", 2 representing "Owner without Mortgage or rent-free", and 3 representing "Renter".
-
-4. spm_totval: household's cash income, unit in USD.
-
-5. spm_snapsub: Supplemental Nutrition Assistance Program (SNAP) subsidy, unit in USD.
-
-6. spm_caphousesub: capped housing subsidy, unit in USD.
-
-7. spm_schlunch: school lunch subsidy, unit in USD.
-
-8. spm_engval: energy subsidy, unit in USD.
-
-9. spm_wicval: Women, Infants, and Children subsidy, unit in USD.
-
-10. spm_fedtax: Federal tax, unit in USD.
-
-11. spm_eitc: Federal Earned Income Tax Credit, unit in USD.
-
-12. spm_actc: Additional Child Tax Credit, unit in USD.
-
-13. spm_fica: Federal Insurance Contributions Act and federal retirement contribution, unit in USD.
-
-14. spm_sttax: state tax, unit in USD.
-
-15. spm_childsuppd: child support paid, unit in USD.
-
-16. spm_capwkccxpns: capped work and child care expense, unit in USD.
-
-17. spm_medxpns: Medical Out-Of-Pocket (MOOP) and Medicare Part B subsidy, unit in USD.
-
 ```{r}
@@
   kable()
+```
+
+```{r}
+#| label: tbl-feature
+#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
+#| warning: false
+
+# show data feature
+feature <- tibble(
+  Data_feature = c("h_seq", "spm_poor", "spm_tenmortstatus", "spm_totval", "spm_snapsub", "spm_caphousesub", "spm_schlunch", "spm_engval", "spm_wicval", "spm_fedtax", "spm_eitc", "spm_actc", "spm_fica", "spm_sttax", "spm_childsuppd", "spm_capwkccxpns", "spm_medxpns"),
+  Data_definition = c("Household sequence number, unique identifier for each household.", "the poverty status; 1 representing 'in poverty' and 0 representing 'not in poverty'.", "household's tenure/mortgage status; 1 representing 'owner with Mortgage', 2 representing 'owner without Mortgage or rent-free', and 3 representing 'renter'.", "household's cash income, unit in USD.", "Supplemental Nutrition Assistance Program (SNAP) subsidy, unit in USD.", "capped housing subsidy, unit in USD.", "school lunch subsidy, unit in USD.", "energy subsidy, unit in USD.", "Women, Infants, and Children subsidy, unit in USD.", "Federal tax, unit in USD.", "Federal Earned Income Tax Credit, unit in USD.", "Additional Child Tax Credit, unit in USD.", "Federal Insurance Contributions Act and federal retirement contribution, unit in USD.", "state tax, unit in USD.", "child support paid, unit in USD.", "capped work and child care expense, unit in USD.", "Medical Out-Of-Pocket (MOOP) and Medicare Part B subsidy, unit in USD.")
+)
+kable(feature, "latex") %>%
+  column_spec(2, width = "30em")
 ```

### 2024-04-02 paper/paper.qmd
add ref

--- 2024-04-01 paper/paper.qmd
@@
 We first chose these 17 variables: h_seq, spm_poor, spm_tenmortstatus, spm_totval, spm_snapsub, spm_caphousesub, spm_schlunch, spm_engval, spm_wicval, spm_fedtax, spm_eitc, spm_actc, spm_fica, spm_sttax, spm_childsuppd, spm_capwkccxpns, and spm_medxpns.
+Their explanation are in @tbl-feature.
 This is because we want to calculate the income after tax, plus all the subsidies and benefits, minus all the expenses, just like how the SPM does it [@sgp2019].
 If we only choose income before tax to analyze, then our analysis will be like the official poverty measure, which requires improvement.
-
-1. h_seq: Household sequence number, unique identifier for each household.
-
-2. spm_poor: the poverty status; 1 representing "In poverty" and 0 represneting "Not in poverty".
-
-3. spm_tenmortstatus: household's tenure/mortgage status; 1 representing "Owner with Mortgage", 2 representing "Owner without Mortgage or rent-free", and 3 representing "Renter".
-
-4. spm_totval: household's cash income, unit in USD.
-
-5. spm_snapsub: Supplemental Nutrition Assistance Program (SNAP) subsidy, unit in USD.
-
-6. spm_caphousesub: capped housing subsidy, unit in USD.
-
-7. spm_schlunch: school lunch subsidy, unit in USD.
-
-8. spm_engval: energy subsidy, unit in USD.
-
-9. spm_wicval: Women, Infants, and Children subsidy, unit in USD.
-
-10. spm_fedtax: Federal tax, unit in USD.
-
-11. spm_eitc: Federal Earned Income Tax Credit, unit in USD.
-
-12. spm_actc: Additional Child Tax Credit, unit in USD.
-
-13. spm_fica: Federal Insurance Contributions Act and federal retirement contribution, unit in USD.
-
-14. spm_sttax: state tax, unit in USD.
-
-15. spm_childsuppd: child support paid, unit in USD.
-
-16. spm_capwkccxpns: capped work and child care expense, unit in USD.
-
-17. spm_medxpns: Medical Out-Of-Pocket (MOOP) and Medicare Part B subsidy, unit in USD.
 
@@
 ## Data Measurement
+The data is mainly collected by surveys.
+Portions of SPM resources, such as values of taxes and some non-cash benefits, are not asked of respondents and are estimated using models.
+The models takes in administrative data where appropriate to ensure that the estimated amounts reflect external totals and distributions [@spm].
 
@@
 ```
-\newpage
 
@@
 
-## mortgage interest rate etc...
+## Limitations
 
-## Weaknesses and next steps
+### Data Consideration
+Firstly, the data are estimates based on a sample of the population, so it cannot represent the whole population of the United States.
+This creates margins of error.
+Also, since the SPM poverty estimates are from household survey data, the data might be affected by issues such as under-reporting, over-reporting, or making up data.
+Some people might not remember the exact value of some data, for example, not all people have or know that they receive income from government benefit programs, so they might be reporting it inaccurately.
+In addition, some SPM resources, such as taxes and non-cash benefits, are estimated using models.
+The models use administrative data, but the estimated amounts are not perfect.
 
-Weaknesses and next steps should also be included.
+### Measurement of Medical Subsidies
+Even the SPM accounts for expenses for healthcare and insurance by subtracting them from family resources, it does not directly value health insurance provided as a resource [@spm].
+The treatment of medical expenses does not take account of all the economic effects of subsidized medical care, for example, there are families who purchase a less comprehensive insurance plan and decide to forego some health care.
+Moreover, some of the largest non-cash benefits programs (Medicare, Medicaid, etc.) are not explicitly taken into account when determining poverty status.
+
+### Lump sum nature of tax credit
+In the SPM, taxes paid and tax credits in the year that are earned are include [@spm]d.
+But typically the credit would not appear in the family income until the next year.
+Also, tax credits are given as a lump sum, but poverty is a something that might happen anytime in the year.
+For example, a family may experience poverty because someone lost a job and cannot find another, so the family is in poverty.
+The tax credit may or may not provide relief depening on the timing in the year and the severity of expenses faced by the family throughout the year.
+So it is not very accurate that the SPM examine resources in a full calendar year compared with a threshold based on the calendar year.
+The family might have experienced 11 months of poverty before they got their tax credits in the 12th month.
+
+## Future Steps
+not done
 
@@
 
+# Data Features
+
+```{r}
+#| label: tbl-feature
+#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
+#| warning: false
+
+# show data feature
+feature <- tibble(
+  Data_feature = c("h_seq", "spm_poor", "spm_tenmortstatus", "spm_totval", "spm_snapsub", "spm_caphousesub", "spm_schlunch", "spm_engval", "spm_wicval", "spm_fedtax", "spm_eitc", "spm_actc", "spm_fica", "spm_sttax", "spm_childsuppd", "spm_capwkccxpns", "spm_medxpns"),
+  Data_definition = c("Household sequence number, unique identifier for each household.", "the poverty status: 1 representing 'in poverty' and 0 representing 'not in poverty'.", "household's tenure/mortgage status: 1 representing 'owner with Mortgage', 2 representing 'owner without Mortgage or rent-free', and 3 representing 'renter'.", "household's cash income, unit in USD.", "Supplemental Nutrition Assistance Program (SNAP) subsidy, unit in USD.", "capped housing subsidy, unit in USD.", "school lunch subsidy, unit in USD.", "energy subsidy, unit in USD.", "Women, Infants, and Children subsidy, unit in USD.", "Federal tax, unit in USD.", "Federal Earned Income Tax Credit, unit in USD.", "Additional Child Tax Credit, unit in USD.", "Federal Insurance Contributions Act and federal retirement contribution, unit in USD.", "state tax, unit in USD.", "child support paid, unit in USD.", "capped work and child care expense, unit in USD.", "Medical Out-Of-Pocket (MOOP) and Medicare Part B subsidy, unit in USD.")
+)
+kable(feature, "latex") %>%
+  column_spec(2, width = "30em")
+```
 # Model details {#sec-model-details}

### 2024-04-11 paper/datasheet_template.qmd
some changes

---
title: "Datasheet for 'US Poverty Status in 2019'"
author: 
  - Mary Cheng
thanks: "Code and data are available at: https://github.com/marycx/us_poverty_analysis_2019.git"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of the United States household poverty status. We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - TBD
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- TBD
2. *How many instances are there in total (of each type, if appropriate)?*
	- TBD
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- TBD
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-04-11 paper/paper.qmd
some changes

@@
 date-format: long
-abstract: "This paper examines household poverty status in the United States in 2019 using the 2019 Revised Supplemental Poverty Measure (SPM) Research Dataset from the United States Census Bureau.
-Using logistic regression analysis, we investigate the influence of household income level after tax and mortgage state on poverty status.
+abstract: "This paper examines household poverty status in the United States in 2019 using the 2019 Revised Supplemental Poverty Measure Research Dataset from the United States Census Bureau.
+Using logistic regression analysis, we investigate the influence of household income level and mortgage state on poverty status.
 Our findings reveal a higher likelihood of poverty for households' total annual earning less than $50000 and for renters compared to property owners.
 All households with incomes exceeding $100000 are not in poverty.
+This implies that household mortgage status and income plays an important role in poverty status.
 These insights emphasize the need for targeted interventions to address socioeconomic inequalities and promote financial well-being among US households."
@@
 # Introduction
-Over the last few decades, addressing poverty has remained a persistent issue in the United States, leading to the implementation of many policies in an effort to mitigate this problem [@Kearney_Harris].
+Over the last few decades, the United States have implemented many policies, in an effort to address and mitigate poverty [@Kearney_Harris].
 In 2019, the United States experienced the culmination of an economic expansion that followed the Great Recession.
 The robustness of the U.S.
-economy during this period was evidenced by low poverty rates, indicating a notable decrease in economic hardship.
-The poverty rate in the United States was estimated to be 10.5% in 2019, down from 11.8% in 2018, marking the fifth consecutive year of decline [@Semega_2022].
-In order to reduce poverty further, it is important for us to analyze why some US household are in poverty in 2019 and how various demographic and financial factors may has an influence on it.
+economy during this period was evidenced by low poverty rates, which was estimated to be 10.5% in 2019, down from 11.8% in 2018, marking the fifth consecutive year of decline [@Semega_2022].
+In order to reduce poverty further, it is important for us to analyze why some US household are in poverty in 2019.
 We analyze the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
@@
 Federal Statistical System, responsible for producing data about the American people and economy [@uscb].
-Supplemental Poverty Measure was developed to address numerous concerns of official poverty measure critics, and its intent is to provide an improved statistical picture of poverty [@Gesumaria_2015].
+Supplemental Poverty Measure was developed to address numerous concerns of official poverty measure critics, providing an improved statistical picture of poverty [@Gesumaria_2015].
 We build a prediction model for households poverty status using 2019 Revised SPM dataset.
-This allows us to discover trends and address potential financial and societal situations that may cause the poverty status for some US households in 2019.
+We primarily look at two factors: mortgage status and income in each household.
+This allows us to look at the question whether potential financial factors may cause the poverty status for some US households in 2019.
 
@@
 Logistic regression is a great choice since it is used to predict binary outcomes, such as poverty status (in poverty or not in poverty).
-Our analysis focuses on estimating the likelihood of household being in poverty, based on various demographic and financial factors captured in the SPM dataset.
+Our analysis focuses on estimating the likelihood of household being in poverty, based on various financial factors captured in the SPM dataset.
 We select 17 data features for analysis.
-After some data cleaning, we summarize 17 features into 2 variables: 1) mortgage state and 2) household annual total income after tax, plus benefits and minus expenses.
+After some data cleaning, we summarize 17 features into 2 variables: 1) mortgage state and 2) household annual total income after tax, plus government benefits and minus expenses.
 The estimand in this paper is the number of households who are in poverty in reality.
 However, it is difficult to measure the exact number of households who are in poverty since there are millions of people in the United States and not all of them will be assessed due to various difficulties.
+For example, due to regional issues, some people who live in rural areas are not included in the survey.
 Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCS.
@@
 Specifically, we assume a normal distribution with a mean of 0 and a standard deviation of 2.5 for each parameter.
+Setting the prior to 0 assumes there is no poverty in any household regardless of their mortgage or income.
+Then, when we factor in real observations, we can see how things shift.
+If the distribution shifts left (less than 0), then we know that particular mortgage or income state is more likely to result in household poverty.
+This approach gives us a clear way to see how data affects the financial situation of households
 
@@
 Also, Bayesian methods enable us to integrate prior knowledge and uncertainty into the analysis, resulting in more reliable estimates of the model parameters.
-
-While alternative modeling approaches like linear regression were considered, we selected Bayesian logistic regression due to the binary nature of our outcome variable, poverty status.
+While alternative modeling approaches like linear regression were considered, we selected Bayesian logistic regression due to the binary nature of our outcome variable: poverty status.
 
@@
 These programs are all aimed at helping families in need to get back on their feet and improve their poverty situation.
+
+## Policy Implications
+Broaden the discussion to include potential policy implications and recommendations based on your findings. How can policymakers use this information to address household poverty?
 

### 2024-04-11 scripts/04-model.R

@@ -26,6 +26,7 @@ analysis_data$poverty_status <- ifelse(analysis_data$poverty_status == "In pover
 # Model 1 for n = 1000
 set.seed(215)
 
+# get the reduced dataset of only 1000 randomly selected data entries
 poverty_data_reduced <- 
   analysis_data |> 
   slice_sample(n = 1000)

### 2024-04-11 paper/datasheet_template.qmd
Datasheet

@@
 thanks: "Code and data are available at: https://github.com/marycx/us_poverty_analysis_2019.git"
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
+date: today
+date-format: long
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
@@
     - The dataset was created to enable analysis of the United States household poverty status.
-We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
+The dataset analyzes poverty status using various demographic and financial factors.
+There is no specific gap that needed to be filled.
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - TBD
+    - United States Census Bureau.
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - TBD
+    - Not sure.
 4. *Any other comments?*
-    - TBD
+    - N/A
 
@@
 If so, please describe how.*
-	- TBD
+	- Majority of the data was acquired in the survey. It was reported by survey participants. Some portion of the data is indirectly derived using model-based guesses.
 2.
@@
 How were these mechanisms or procedures validated?*
-	- TBD
+	- Surveys and software programs for the model.
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- TBD
+	- Not sure.
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- TBD
+	- United States Census Bureau workers.
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- TBD
+	- Data was collected in some time in 2019.
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- Not sure
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- TBD
+	- Surveys
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- TBD
+	- Not sure
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- TBD
+	- Not sure
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- TBD
+	- Not sure
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- Not sure
 12. *Any other comments?*
-	- TBD
+	- N/A
 
@@
 If not, you may skip the remaining questions in this section.*
-	- TBD
+	- Yes.
+The original dataset includes duplicate responses from the same household.
+As the dataset features are related to the entire household as a whole, this paper will eliminate duplicate entries, using the variable "h_seq" which is an unique household identifier, to ensure data accuracy and consistency.
+First, a new column "income after tax" is created by summing up all incomes, credits, and benefits and subtracting all expenses.
+Then the dataset is cleaned by renaming of column headers, defining column classes, grouping the income variable into different income levels (here the assumed facts are that the income unit is in USD and 'k' stands for thousand, eg.
+5k = 10,000), and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- TBD
+	- Yes. https://github.com/marycx/us_poverty_analysis_2019/tree/main/data/raw_data
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
+	- R language 
 4. *Any other comments?*
-	- TBD
+	- N/A
 
@@
 1. *Has the dataset been used for any tasks already? If so, please provide a description.*
-	- TBD
-2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- TBD
+	- Yes, for 2019 poverty status analysis by US government.
+2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other acess point.*
+	- No.
 3. *What (other) tasks could the dataset be used for?*
-	- TBD
+	- The dataset can also be analyzed to show relationship between different data features and see how those relate to each other.
 4.
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
+	- No.
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- No.
 6. *Any other comments?*
-	- TBD
+	- N/A
 
@@
 If so, please provide a description.*
-	- TBD
+	- No.
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- TBD
+	- Github. No DOI for now.
 3. *When will the dataset be distributed?*
-	- TBD
+	- 2024. 04. 10
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- TBD
+	- No.
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- TBD
+	- No.
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
+	- No.
 7. *Any other comments?*
-	- TBD
+	- N/A
 
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- TBD
+	- Mary Cheng
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- TBD
+	- maryc.cheng@mail.utoronto.ca
 3. *Is there an erratum? If so, please provide a link or other access point.*
-	- TBD
+	- No.
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- TBD
+	- No.
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- TBD
+	- No.
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- TBD
+	- No.
 7.
@@
 If so, please provide a description.*
-	- TBD
+	- Yes, open an issue on my Github repository.
 8. *Any other comments?*
-	- TBD
+	- N/A
 
 \newpage
-
-# References

### 2024-04-12 paper/datasheet_template.qmd
datasheet update

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
 format: pdf
 number-sections: true
-bibliography: references.bib
 ---
@@
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - United States Census Bureau.
+    - Mary Cheng for STA302 course at University of Toronto
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
-    - Not sure.
+    - No direct funding was received for this project.
 4. *Any other comments?*
@@
 Please provide a description.*
-	- TBD
+	- Each row of the main dataset represents a household.
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- TBD
+	- There are 62990 instances.
 3.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-	- TBD
+	- The dataset is a sample.
+The larger set contains duplicate data entries from the same household.
+As the features of the dataset are related to the entire household as a whole, I decided to remove all the duplicated entries.
+The sample is representative of the larger set because I kept all the unique data entries.
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
-	- TBD
+	- Each instance consists of financial information such as household mortgage state, income after tax plus all the benefits and minus all the expenses.
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- TBD
+	- There is a unique key associated with each household.
 6.
@@
 This does not include intentionally removed information, but might include, for example, redacted text.*
-	- TBD
+	- No.
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- TBD
+	- Yes, through the unique ID.
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- TBD
+	- No.
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- TBD
+	- There is some human uncertainties in the dataset, for example, respondents might report false information in the survey.
 10.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- TBD
+	- Self-contained
 11.
@@
 If so, please provide a description.*
-	- TBD
+	- No.
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	- TBD
+	- No.
 13.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- TBD
+	- No.
 14.
@@
 If so, please describe how.*
-	- TBD
+	- Maybe. If you include other information like government benefit programs with this information, you might be able to cross reference some individuals.
 15.
@@
 If so, please provide a description.*
-	- TBD
+	- Yes, financial information is contained in this dataset.
 16. *Any other comments?*
-	- TBD
+	- No.
 
@@
 If so, please describe how.*
-	- Majority of the data was acquired in the survey. It was reported by survey participants. Some portion of the data is indirectly derived using model-based guesses.
+	- The data was acquired by downloading 2019 Revised Supplemental Poverty Measure (SPM) Research dataset provided by the United States Census Bureau.
 2.
@@
 How were these mechanisms or procedures validated?*
-	- Surveys and software programs for the model.
+	- Download the data using R
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- Not sure.
+	- It is not a sample.
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- United States Census Bureau workers.
+	- Mary Cheng
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- Data was collected in some time in 2019.
+	- Data was collected in April 2024. The data was first created by the United States Census Bureau in 2019.
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- Not sure
+	- No.
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- Surveys
+	- Third parties.
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- Not sure
+	- No.
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- Not sure
+	- No.
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- Not sure
+	- N/A
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- Not sure
+	- No.
 12. *Any other comments?*
@@
 	- Yes.
+Cleaning data is done.
 The original dataset includes duplicate responses from the same household.
@@
 If so, please provide a description.*
-	- No.
+	- The dataset is available through Github.
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
@@
 3. *When will the dataset be distributed?*
-	- 2024. 04. 10
+	- It is available now.
 4.
@@
 If so, please provide a description.*
-	- Yes, open an issue on my Github repository.
+	- Pull request on Github.
 8. *Any other comments?*

### 2024-04-12 paper/paper.qmd
datasheet update

@@
 ## Data Measurement
-The data is mainly collected by surveys.
-Portions of SPM resources, such as values of taxes and some non-cash benefits, are not asked of respondents and are estimated using models.
-The models takes in administrative data where appropriate to ensure that the estimated amounts reflect external totals and distributions [@spm].
+The data was mainly collected by surveys.
+Portions of SPM resources, such as values of taxes and some non-cash benefits, were not asked of respondents and were estimated using models.
+However, the estimated amounts are not perfect.
+For example, they found that the estimated total benefits from both the Earned Income Tax Credit and the child tax credit are lower than those found when examining federal income tax returns [@spm].
 
 ## Methodology
-
-The original dataset includes duplicate responses from the same household.
-As the dataset features are related to the entire household as a whole, this paper will eliminate duplicate entries, using the variable "h_seq" which is an unique household identifier, to ensure data accuracy and consistency.
+The original dataset includes duplicated responses from the same household.
+As the report is analyzing the household poverty status, I will eliminate duplicate entries, using the variable "h_seq" which is an unique household identifier, to ensure data accuracy and consistency.
 

### 2024-04-16 paper/paper.qmd
edit according to TA review

@@
 In 2019, the United States experienced the culmination of an economic expansion that followed the Great Recession.
-The robustness of the U.S.
-economy during this period was evidenced by low poverty rates, which was estimated to be 10.5% in 2019, down from 11.8% in 2018, marking the fifth consecutive year of decline [@Semega_2022].
+The poverty rate was estimated to be 10.5% in 2019 [@Semega_2022].
 In order to reduce poverty further, it is important for us to analyze why some US household are in poverty in 2019.
@@
 After some data cleaning, we summarize 17 features into 2 variables: 1) mortgage state and 2) household annual total income after tax, plus government benefits and minus expenses.
-The estimand in this paper is the number of households who are in poverty in reality.
-However, it is difficult to measure the exact number of households who are in poverty since there are millions of people in the United States and not all of them will be assessed due to various difficulties.
-For example, due to regional issues, some people who live in rural areas are not included in the survey.
-Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCS.
 
@@
 Statistical programming language R [@citeR] is used in this report, with packages `tidyverse` [@citeTidyverse], `here` [@citeHere], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModleSummary], `ggplot2` [@citeGgplot2], `knitr` [@citeKnitr], `marginaleffects` [@citeMarginalEffects], `plotly` [@citePlotly], `tibble` [@citeTibble], `margins` [@citeMargins], `testthat` [@citetestthat] and `kableExtra` [@citeKableExtra].
+
+## Estimand
+The estimand in this paper is the number of households who are in poverty in reality.
+However, it is difficult to measure the exact number of households who are in poverty since there are millions of people in the United States and not all of them will be assessed due to various difficulties.
+For example, due to regional issues, some people who live in rural areas are not included in the survey.
+Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is fitted using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCS.
 
@@
 ## Features
-The original SPM 2019 dataset, which shows in @tbl-raw, contains 157959 data entries and many variables.
+The original SPM 2019 dataset, which shows in @tbl-raw in Appendix @sec-data-details, contains 157959 data entries and many variables.
 Since it is difficult to observe such a large dataset, this report will only explore and analyze through several data features.
@@
 Their explanation are in @tbl-feature.
-This is because we want to calculate the income after tax, plus all the subsidies and benefits, minus all the expenses, just like how the SPM does it [@sgp2019].
-If we only choose income before tax to analyze, then our analysis will be like the official poverty measure, which requires improvement.
+We can use those features (other than h_seq, spm_poor, spm_tenmortstatus) to calculate the income after tax, plus all the subsidies and benefits, minus all the expenses, just like how the SPM does it [@sgp2019].
+This way the income is able to reflect the real financial situation on households more accurately.
 
@@
 
-```{r}
-#| label: tbl-raw
-#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
-#| message: false
-#| echo: false
-
-raw_poverty_data[1:5, 1:6] |>
-  kable()
-raw_poverty_data[1:5, 7:13] |>
-  kable()
-raw_poverty_data[1:5, 13:17] |>
-  kable()
-```
-
 ## Data Measurement
@@
 For example, they found that the estimated total benefits from both the Earned Income Tax Credit and the child tax credit are lower than those found when examining federal income tax returns [@spm].
+
+The unit of measurement for numerical data are in US Dollars. For other data, they are categorical features with no units.
 
@@
 5k = 10,000), and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
-After cleaning, 62990 rows of data with 3 data features remain.
-@tbl-clean-data shows a preview of the cleaned dataset.
+
+After cleaning, 62990 rows of data with 3 data features remain. @tbl-clean-data shows a preview of the cleaned dataset.
 
@@
 
-```{r}
+```{r fig.pos="h"}
 #| label: tbl-clean-data
@@
 #| echo: false
+
+colnames(poverty2019) <- c("poverty status", "mortgage state", "income level")
 
@@
 
-```{r}
+```{r fig.pos="h"}
 #| label: tbl-summary
@@
 
-```{r}
+```{r fig.pos="h"}
 #| label: fig-mortgage
@@
 #| message: false
+
+colnames(poverty2019) <- c("poverty_status", "mortgage_state", "income")
 poverty2019 |>
@@
 
-```{r}
+```{r fig.pos="h"}
 #| label: fig-income
@@
 
-## Model Implication
-For posterior predictive checks, in @fig-post_dist, the great fit of the posterior distribution from our logistic regression model with actual poverty data suggests accurate capture of poverty status patterns.
-This indicates the accuracy of our model's poverty status prediction to the 2019 SPM data.
-Also, @fig-post_prior compares the posterior to the prior, which shows some parameter changes such as "below 5k" and "5k-50k", as well as the intercept (people who are owners with mortgage and earn 100-150k).
-The discrepancy for the parameter suggest that our prior may not be very accurate in regards to these specific aspects.
-
-The trace plots in @fig-trace1 and @fig-trace2 do not suggest anything out of the ordinary.
-Also, with the Rhat plot (@fig-rhat), we can observe that everything is close to 1, and no more than 1.05, which shows the great convergence of Markov chain Monte Carlo for our model.
-
-More detailed explanation of each plot can be found in Appendix @sec-model-details.
-
 # Results {#sec-result}
@@
 
-```{r}
+## Model Validation
+For posterior predictive checks, in @fig-post_dist, the great fit of the posterior distribution from our logistic regression model with actual poverty data suggests accurate capture of poverty status patterns.
+This indicates the accuracy of our model's poverty status prediction to the 2019 SPM data.
+Also, @fig-post_prior compares the posterior to the prior, which shows some parameter changes such as "below 5k" and "5k-50k", as well as the intercept (people who are owners with mortgage and earn 100-150k).
+The discrepancy for the parameter suggest that our prior may not be very accurate in regards to these specific aspects.
+
+The trace plots in @fig-trace1 and @fig-trace2 do not suggest anything out of the ordinary.
+Also, with the Rhat plot (@fig-rhat), we can observe that everything is close to 1, and no more than 1.05, which shows the great convergence of Markov chain Monte Carlo for our model.
+
+More explanation can be found in Appendix @sec-model-details.
+```{r fig.pos="h"}
 #| echo: false
@@
 
+\newpage
 # Discussion {#sec-discussion}
@@
 
-## Policy Implications
-Broaden the discussion to include potential policy implications and recommendations based on your findings. How can policymakers use this information to address household poverty?
-
 ## Limitations
@@
 For example, a family may experience poverty because someone lost a job and cannot find another, so the family is in poverty.
-The tax credit may or may not provide relief depening on the timing in the year and the severity of expenses faced by the family throughout the year.
+The tax credit may or may not provide relief depending on the timing in the year and the severity of expenses faced by the family throughout the year.
 So it is not very accurate that the SPM examine resources in a full calendar year compared with a threshold based on the calendar year.
@@
 ## Future Steps
-not done
+Future research should focus on gathering more inclusive and representative data by increasing the volume of data entries and mitigating bias in data collection.
+Beyond simply collecting data about household financial data, researchers should investigate more government benefits and expenses that are imposed on households.
+By investigating those benefits and expenses, researchers can arrive at conclusions if the current government poverty policies are effective for household in poverty.
+It is important to ensure that households in poverty have access to the necessary resources and support by the US government.
+This makes sure that the research not only analyzes the situation, but also contributes to actionable improvements for those affected.
 
@@
 
-# Data Features
+# Data {#sec-data-details}
+## Raw data
+@tbl-raw shows the preview of the raw 2019 US Supplemental Poverty Measure dataset.
+```{r fig.pos="h"}
+#| label: tbl-raw
+#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
+#| message: false
+#| echo: false
 
-```{r}
+raw_poverty_data[1:5, 1:6] |>
+  kable()
+raw_poverty_data[1:5, 7:13] |>
+  kable()
+raw_poverty_data[1:5, 13:17] |>
+  kable()
+```
+
+## Data feature
+@tbl-feature shows the selected features of the 2019 US Supplemental Poverty Measure dataset.
+```{r fig.pos="h"}
 #| label: tbl-feature
-#| tbl-cap: "Preview of the raw 2019 US Supplemental Poverty Measure dataset"
+#| tbl-cap: "Selected feature of the 2019 US Supplemental Poverty Measure dataset"
 #| message: false
@@
 ## 90% Credibility Interval {#sec-credibility-interval}
+@fig-modelresults1 and @fig-modelresults2 are 90% credibility interval plots.
 ```{r}

### 2024-04-17 paper/paper.qmd
some edit

@@
 # Introduction
-Over the last few decades, the United States have implemented many policies, in an effort to address and mitigate poverty [@Kearney_Harris].
-In 2019, the United States experienced the culmination of an economic expansion that followed the Great Recession.
-The poverty rate was estimated to be 10.5% in 2019 [@Semega_2022].
-In order to reduce poverty further, it is important for us to analyze why some US household are in poverty in 2019.
-We analyze the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
+Over the last few decades, despite many policies the United States have implemented to mitigate poverty, there has been no real improvement [@Kearney_Harris].
+Poverty remains a major issue in the US, affecting millions of people.
+In 2019, the poverty rate was 10.5% [@Semega_2022].
+Therefore, it is important to find out what are the causes of poverty, in order to reduce it.
+Multiple research, using the Official Poverty Measure, was conducted to look at why poverty persists.
+However, the Official Poverty Measure is flawed in various aspects, such as not accounting for regional differences in costs of living and government benefits  [@Gesumaria_2015].
+So we will analyze why some US household are in poverty in 2019 using the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
 USCB is an agency of the U.S.
 Federal Statistical System, responsible for producing data about the American people and economy [@uscb].
-Supplemental Poverty Measure was developed to address numerous concerns of official poverty measure critics, providing an improved statistical picture of poverty [@Gesumaria_2015].
+Supplemental Poverty Measure was developed to address numerous concerns of Official Poverty Measure critics, providing an improved statistical picture of poverty [@Gesumaria_2015].
 We build a prediction model for households poverty status using 2019 Revised SPM dataset.
 We primarily look at two factors: mortgage status and income in each household.
-This allows us to look at the question whether potential financial factors may cause the poverty status for some US households in 2019.
+This allows us to look at the question whether potential financial factors may cause the poverty status for US households in 2019.
 
@@
 Specifically, we assume a normal distribution with a mean of 0 and a standard deviation of 2.5 for each parameter.
-Setting the prior to 0 assumes there is no poverty in any household regardless of their mortgage or income.
-Then, when we factor in real observations, we can see how things shift.
-If the distribution shifts left (less than 0), then we know that particular mortgage or income state is more likely to result in household poverty.
-This approach gives us a clear way to see how data affects the financial situation of households
 

### 2024-04-17 paper/paper.qmd
based on TA review

@@
 # Introduction
-Over the last few decades, despite many policies the United States have implemented to mitigate poverty, there has been no real improvement [@Kearney_Harris].
+Despite many policies the United States have implemented to mitigate poverty over the past decades, there has been no real improvement [@Kearney_Harris].
 Poverty remains a major issue in the US, affecting millions of people.
 In 2019, the poverty rate was 10.5% [@Semega_2022].
-Therefore, it is important to find out what are the causes of poverty, in order to reduce it.
+Therefore, it is important to find out the causes of poverty, in order to reduce it.
 Multiple research, using the Official Poverty Measure, was conducted to look at why poverty persists.
@@
 We primarily look at two factors: mortgage status and income in each household.
-This allows us to look at the question whether potential financial factors may cause the poverty status for US households in 2019.
+This allows us to look at the question whether potential financial factors may influence the poverty status for US households in 2019.
 
@@
 Then the dataset is cleaned by renaming of column headers, defining column classes, grouping the income variable into different income levels (here the assumed facts are that the income unit is in USD and 'k' stands for thousand, eg.
-5k = 10,000), and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
+5k = 5000), and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
 
@@
 
-### Model justification
+## Model justification
 Regarding the relationship between mortgage state and poverty state, we anticipate that both owners without mortgages and owners with mortgages are less likely to experience poverty.
@@
 In the USA in 2019, housing costs accounted for more than 30% of household expenditure, making mortgage status a critical factor in determining a family's risk of falling into poverty [@Statistics_2020].
-With 11.8% of the total population (38.3 million people) living below the poverty line, it's evident that housing affordability significantly impacts economic stability.
-The percentage of individuals below the poverty threshold, adjusted for housing costs are: $29,080 for a family of four with a mortgage, $24,413 for a family of four without a mortgage, and $29,194 for a family of four [@Khattar_Pathak_Schweitzer_Khan_McManus_Jones_Benson_Kelly_2023].
-From this, we can assumed that families without mortgages experienced the lowest housing costs, leaving more funds available for other essentials and eliminating the need for alternative housing options.
+The percentage of individuals below the poverty threshold, adjusted for housing costs are: $29,080 for a family of four with a mortgage, $24,413 for a family of four without a mortgage, and $29,194 for a family of four renting [@Khattar_Pathak_Schweitzer_Khan_McManus_Jones_Benson_Kelly_2023].
+From this, we can assumed that families without mortgages experienced the lowest housing costs, leaving more funds available for other essentials.
 In contrast, families with mortgages may face burdensome monthly payments, especially if they have chosen a floating interest rate, making them vulnerable to fluctuations in interest rates.
 Renters also encounter challenges, as rent increases can strain finances and necessitate either accepting higher housing costs or undertaking the time-consuming and costly process of finding new accommodations.
-These housing-related challenges exacerbate the risk of poverty for affected families, underscoring the importance of addressing housing affordability as a key component of poverty alleviation efforts.
+
+Also, low-income households face challenges with accessing lower mortgage rates and reducing debt burdens through refinancing.
+Households with low to moderate incomes were less able to take advantage of lower mortgage rates and alleviate their debt burdens through refinancing during the period of exceptionally low rates: merely 24% of mortgages in low-income regions underwent refinancing, in contrast to the 42% in high-income areas [@mortgage].
+Therefore, the low-income households have harder experience with repaying their mortgage than higher income households.
+
+These housing-related challenges increases the risk of poverty for affected families, underscoring the importance of addressing housing affordability as a key component of poverty alleviation efforts.
 
@@
 However, in the United States, the cost of living is notably high, low-income households often struggle to meet basic needs, resulting in an unending cycle of poverty.
+About half of lower-income parents (52%) say they have not had enough money for food or their rent or mortgage [@Braga_2022].
+Some 45% of employed parents with lower incomes say they would be very worried about losing pay if they needed to take a day or two off from work to deal with a sick child or because of child care issues.
+This type of delimma for low-income households makes them harder to pull themselves out of poverty.
+
 In 2019, the poverty guideline for a one-person household was set at $12,490 annually [@ASPE].

### 2024-04-18 paper/datasheet_2019SPM.qmd
rename

--- 2024-04-12 paper/datasheet_template.qmd
@@
 ---
-title: "Datasheet for 'US Poverty Status in 2019'"
+title: "Datasheet for '2019 Revised Supplemental Poverty Measure Research Dataset'"
 author: 
@@
 number-sections: true
+bibliography: references.bib
 ---
@@
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - Mary Cheng for STA302 course at University of Toronto
+    - United States Census Bureau
 3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
@@
 Please provide a description.*
-	- Each row of the main dataset represents a household.
+	- Each row of the main dataset represents an individual in a household.
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- There are 62990 instances.
+	- There are 157,959 instances.
 3.
@@
 	- The dataset is a sample.
-The larger set contains duplicate data entries from the same household.
-As the features of the dataset are related to the entire household as a whole, I decided to remove all the duplicated entries.
-The sample is representative of the larger set because I kept all the unique data entries.
+The larger set should be all the people in the United States of America.
+The sample is not fully representative of the larger set.
+Even though the sample takes various demographic variables into consideration, there are still people in the US who are not surveyed.
+This creates a potential problem that some people with different age, race, and background might not be accounted for.
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
-	- Each instance consists of financial information such as household mortgage state, income after tax plus all the benefits and minus all the expenses.
+	- Each instance consists of different demographic and financial information of each participant in this survey.
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- There is a unique key associated with each household.
+	- No. There is unique identifier for each household, however, if there are more than one person in the household, these people will have the same identifier since they live in the same household.
 6.
@@
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- Yes, through the unique ID.
+	- No, since there is no unique identifier associated with each individual instance.
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- Self-contained
+	- Self-contained.
 11.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- No.
+	- Yes, age and race. 
 14.
@@
 If so, please describe how.*
-	- The data was acquired by downloading 2019 Revised Supplemental Poverty Measure (SPM) Research dataset provided by the United States Census Bureau.
+	- The data was mainly collected by surveys by United States Census Bureau.
+Portions of SPM resources, such as values of taxes and some non-cash benefits, were not asked of respondents and were estimated using models [@spm].
 2.
@@
 How were these mechanisms or procedures validated?*
-	- Download the data using R
+	- Surveys and a computer model.
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
@@
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- Mary Cheng
+	- workers from United States Census Bureau.
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- Data was collected in April 2024. The data was first created by the United States Census Bureau in 2019.
+	- The data was created by the United States Census Bureau in 2019.
 6.
@@
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- Third parties.
+	- Not sure.
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- No.
+	- Yes, it was done by survey [@spm].
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- No.
+	- Not sure.
 10.
@@
 If not, you may skip the remaining questions in this section.*
-	- Yes.
-Cleaning data is done.
-The original dataset includes duplicate responses from the same household.
-As the dataset features are related to the entire household as a whole, this paper will eliminate duplicate entries, using the variable "h_seq" which is an unique household identifier, to ensure data accuracy and consistency.
-First, a new column "income after tax" is created by summing up all incomes, credits, and benefits and subtracting all expenses.
-Then the dataset is cleaned by renaming of column headers, defining column classes, grouping the income variable into different income levels (here the assumed facts are that the income unit is in USD and 'k' stands for thousand, eg.
-5k = 10,000), and replacing numerical values in the table with their corresponding descriptions from the data dictionary to improve the readability.
+	- No.
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- Yes. https://github.com/marycx/us_poverty_analysis_2019/tree/main/data/raw_data
+	- N/A
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- R language 
+	- N/A
 4. *Any other comments?*
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- No.
+	- There are some financial information in there that might affect individuals. For example, if the bank knows that you were in poverty for years, they are less likely to give you a loan.
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
@@
 If so, please provide a description.*
-	- The dataset is available through Github.
+	- The dataset is available through United States Census Bureau official website.
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- Github. No DOI for now.
+	- STATA / SAS data download.
 3. *When will the dataset be distributed?*
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- Mary Cheng
+	- No one.
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- maryc.cheng@mail.utoronto.ca
+	- N/A
 3. *Is there an erratum? If so, please provide a link or other access point.*
@@
 If so, please provide a description.*
-	- Pull request on Github.
+	- The dataset is on government website. You may contact the United States Census Bureau officials in ways listed on this website: https://www.census.gov/about/contact-us.html
 8. *Any other comments?*
@@
 \newpage
+**Reference**

### 2024-04-18 paper/US_Poverty_Analysis_2019.qmd
edit grammar

--- 2024-04-17 paper/paper.qmd
@@
 Poverty remains a major issue in the US, affecting millions of people.
-In 2019, the poverty rate was 10.5% [@Semega_2022].
-Therefore, it is important to find out the causes of poverty, in order to reduce it.
-Multiple research, using the Official Poverty Measure, was conducted to look at why poverty persists.
-However, the Official Poverty Measure is flawed in various aspects, such as not accounting for regional differences in costs of living and government benefits  [@Gesumaria_2015].
-So we will analyze why some US household are in poverty in 2019 using the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
+In 2019, the poverty rate was 10.5%, which was approximately 34.0 million people [@Semega_2022].
+Therefore, we want to find out the causes of poverty, in order to reduce it.
+Multiple research, using the Official Poverty Measure dataset gathered by the United States Census Bureau (USCB), looked at why poverty persists.
 USCB is an agency of the U.S.
 Federal Statistical System, responsible for producing data about the American people and economy  [@uscb].
-Supplemental Poverty Measure was developed to address numerous concerns of Official Poverty Measure critics, providing an improved statistical picture of poverty [@Gesumaria_2015].
+Over the years, people have realized that the Official Poverty Measure is flawed in various aspects, such as not accounting for regional differences in costs of living, not considering household's mortgage state, and government benefits programs  [@Gesumaria_2015].
+Therefore, Supplemental Poverty Measure was developed to address numerous concerns of Official Poverty Measure critics, providing an improved statistical picture of poverty .
+In this paper, we will analyze why some US household are in poverty in 2019 using the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
 We build a prediction model for households poverty status using 2019 Revised SPM dataset.
-We primarily look at two factors: mortgage status and income in each household.
 This allows us to look at the question whether potential financial factors may influence the poverty status for US households in 2019.
@@
 Our analysis focuses on estimating the likelihood of household being in poverty, based on various financial factors captured in the SPM dataset.
-We select 17 data features for analysis.
-After some data cleaning, we summarize 17 features into 2 variables: 1) mortgage state and 2) household annual total income after tax, plus government benefits and minus expenses.
+I selected 17 data features for analysis.
+After some data cleaning, I summarized 17 features into 2 variables: 1) mortgage state and 2) household annual total income after tax, plus government benefits and minus expenses.
 
@@
 Therefore, any effects of tax credits designed to assist persons with low income are not considered by the official measure.
-The Supplemental Poverty Measure (SPM), similar to the official poverty measure, tackles the shortcomings of poverty assessment.
-It determines poverty status by gauging resources against a defined standard of living.
-This standard considers expenditures on essentials like food, clothing, shelter, and utilities, plus a bit more for additional expenses.
+So, the Supplemental Poverty Measure (SPM) was developed.
+It is similar to the official poverty measure, but it is more comprehensive.
+It considers expenditures on essentials like food, clothing, shelter, and utilities, plus a bit more for additional expenses.
 The resources assessed include disposable income, accounting for taxes and some non-cash benefits, available to cover these needs [@Gesumaria_2015].

### 2024-04-18 paper/US_Poverty_Analysis_2019.qmd
fixed reference, final edits

@@
 Using logistic regression analysis, we investigate the influence of household income level and mortgage state on poverty status.
-Our findings reveal a higher likelihood of poverty for households' total annual earning less than $50000 and for renters compared to property owners.
-All households with incomes exceeding $100000 are not in poverty.
-This implies that household mortgage status and income plays an important role in poverty status.
+Our findings reveal a higher likelihood of poverty for households' total annual income less than $50000, while all households with incomes exceeding $100000 are not in poverty.
+Also, renters is more likely to be in poverty compared to property owners.
 These insights emphasize the need for targeted interventions to address socioeconomic inequalities and promote financial well-being among US households."
@@
 # Introduction
-Despite many policies the United States have implemented to mitigate poverty over the past decades, there has been no real improvement [@Kearney_Harris].
+Despite many policies the United States have implemented to mitigate poverty, there has been no real improvement [@Kearney_Harris].
 Poverty remains a major issue in the US, affecting millions of people.
@@
 Over the years, people have realized that the Official Poverty Measure is flawed in various aspects, such as not accounting for regional differences in costs of living, not considering household's mortgage state, and government benefits programs  [@Gesumaria_2015].
-Therefore, Supplemental Poverty Measure was developed to address numerous concerns of Official Poverty Measure critics, providing an improved statistical picture of poverty .
+Therefore, Supplemental Poverty Measure was developed to address numerous concerns of Official Poverty Measure critics, providing an improved statistical picture of poverty.
 In this paper, we will analyze why some US household are in poverty in 2019 using the 2019 Revised Supplemental Poverty Measure (SPM) Research dataset from the United States Census Bureau (USCB) [@Bureau_2022].
@@
 However, it is difficult to measure the exact number of households who are in poverty since there are millions of people in the United States and not all of them will be assessed due to various difficulties.
-For example, due to regional issues, some people who live in rural areas are not included in the survey.
+For example, due to regional issues, some people who live in rural areas might not be included in the survey.
 Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is fitted using sample dataset from the 2019 Revised Supplemental Poverty Measure Research Data from USCS.
@@
 For instance, it uses money income before taxes, so it does not measure the income available for individuals to spend, which for most people is after-tax income.
-Therefore, any effects of tax credits designed to assist persons with low income are not considered by the official measure.
 So, the Supplemental Poverty Measure (SPM) was developed.
-It is similar to the official poverty measure, but it is more comprehensive.
+It is similar to the official poverty measure, but it takes more aspects into account.
 It considers expenditures on essentials like food, clothing, shelter, and utilities, plus a bit more for additional expenses.
@@
 Their explanation are in @tbl-feature.
-We can use those features (other than h_seq, spm_poor, spm_tenmortstatus) to calculate the income after tax, plus all the subsidies and benefits, minus all the expenses, just like how the SPM does it [@sgp2019].
-This way the income is able to reflect the real financial situation on households more accurately.
+We can use those features (other than h_seq, spm_poor, spm_tenmortstatus) to calculate the income after tax, by adding all the subsidies and benefits and substracting all the expenses (just like how the SPM does it [@sgp2019]).
+This way the income is able to reflect the financial situation on households more accurately.
 
@@
 ## Data Measurement
-The data was mainly collected by surveys.
+The data was mainly collected by surveys conducted by USCB.
 Portions of SPM resources, such as values of taxes and some non-cash benefits, were not asked of respondents and were estimated using models.
@@
 The unit of measurement for numerical data are in US Dollars. For other data, they are categorical features with no units.
+
+### Data Consideration
+The data are estimates based on a sample of the population, so it cannot represent the whole population of the United States.
+This creates margins of error.
+Also, since the SPM poverty estimates are from household survey data, the data might be affected by issues such as under-reporting, over-reporting, or making up data.
+Some people might not remember the exact value of some data, for example, not all people have or know that they receive income from government benefit programs, so they might be reporting it inaccurately.
+In addition, some SPM resources, such as taxes and non-cash benefits, are estimated using models.
+The models use administrative data, but the estimated amounts are not perfect.
 
@@
 ## Data Visualization
-@fig-mortgage illustrates the relationship between mortgage state and poverty status of households.
+@tbl-mortgage illustrates the relationship between mortgage state and poverty status of households.
 It can be seen that renters are the highest proportion of households experiencing poverty, followed by owners without mortgages, and finally owners with mortgages.
@@
 Despite variations in poverty rates across different mortgage states, the cumulative data shows economic stability among the majority of households in the United States in 2019.
+```{r}
+#| label: tbl-mortgage
+#| tbl-cap: "The distribution of poverty status by mortgage state"
+#| message: false
+#| echo: false
+#| warning: false
+colnames(poverty2019) <- c("poverty_status", "mortgage_state", "income")
 
-```{r fig.pos="h"}
+# calculate poverty ratio for renter, Owner with Mortgage, and Owner without Mortgage
+rent_p <- sum(poverty2019$mortgage_state == "Renter" & poverty2019$poverty_status == "In poverty")
+rent_np <- sum(poverty2019$mortgage_state == "Renter" & poverty2019$poverty_status == "Not in poverty")
+
+mort_p <- sum(poverty2019$mortgage_state == "Owner with Mortgage" & poverty2019$poverty_status == "In poverty")
+mort_np <- sum(poverty2019$mortgage_state == "Owner with Mortgage" & poverty2019$poverty_status == "Not in poverty")
+
+nomort_p <- sum(poverty2019$mortgage_state == "Owner without Mortgage" & poverty2019$poverty_status == "In poverty")
+nomort_np <- sum(poverty2019$mortgage_state == "Owner without Mortgage" & poverty2019$poverty_status == "Not in poverty")
+
+# show data feature
+mortgage <- tibble(
+  Household_type = c("Renters", "Owners with Mortgage", "Owners without Mortgage"),
+  In_poverty_percentage = c(round(rent_p/(rent_p+rent_np)*100, digits = 2), round(mort_p/(mort_p+mort_np)*100, digits = 2), round(nomort_p/(nomort_p+nomort_np)*100, digits = 2)),
+  Not_in_poverty_percentage = c(round(rent_np/(rent_p+rent_np)*100, digits = 2), round(mort_np/(mort_p+mort_np)*100, digits = 2), round(nomort_np/(nomort_p+nomort_np)*100, digits = 2))
+)
+
+kable(mortgage, 
+      align = c("c", "c", "c")  # Center alignment for all columns
+)
+```
+
+```{r fig.show='hide'}
 #| label: fig-mortgage
@@
 #| message: false
+#|
 
@@
 
-@fig-income shows the relationship between household annual total income and their poverty status.
+@tbl-income shows the relationship between household annual total income and their poverty status.
 We can see that among households earning more than 100k annually, no household is in poverty.
@@
 
-```{r fig.pos="h"}
+```{r}
+#| label: tbl-income
+#| tbl-cap: "The distribution of poverty status by income level"
+#| message: false
+#| echo: false
+#| warning: false
+
+# calculate poverty ratio for different income levels
+five_fifty_p <- sum(poverty2019$income == "5k-50k" & poverty2019$poverty_status == "In poverty")
+five_fifty_np <- sum(poverty2019$income == "5k-50k" & poverty2019$poverty_status == "Not in poverty")
+
+belowfive_p <- sum(poverty2019$income == "below 5k" & poverty2019$poverty_status == "In poverty")
+belowfive_np <- sum(poverty2019$income == "below 5k" & poverty2019$poverty_status == "Not in poverty")
+
+fifty_hund_p <- sum(poverty2019$income == "50k-100k" & poverty2019$poverty_status == "In poverty")
+fifty_hund_np <- sum(poverty2019$income == "50k-100k" & poverty2019$poverty_status == "Not in poverty")
+
+hund_fifty_p <- sum(poverty2019$income == "100k-150k" & poverty2019$poverty_status == "In poverty")
+hund_fifty_np <- sum(poverty2019$income == "100k-150k" & poverty2019$poverty_status == "Not in poverty")
+
+fifty_twohund_p <- sum(poverty2019$income == "150k-200k" & poverty2019$poverty_status == "In poverty")
+fifty_twohund_np <- sum(poverty2019$income == "150k-200k" & poverty2019$poverty_status == "Not in poverty")
+
+twohund_fifty_p <- sum(poverty2019$income == "200k-250k" & poverty2019$poverty_status == "In poverty")
+twohund_fifty_np <- sum(poverty2019$income == "200k-250k" & poverty2019$poverty_status == "Not in poverty")
+
+above_p <- sum(poverty2019$income == "above 250k" & poverty2019$poverty_status == "In poverty")
+above_np <- sum(poverty2019$income == "above 250k" & poverty2019$poverty_status == "Not in poverty")
+
+# show data feature
+income_level <- tibble(
+  Household_type = c("below 5k","5k-50k","50k-100k","100k-150k","150k-200k","200k-250k","above 250k"),
+  In_poverty_percentage = c(round(belowfive_p/(belowfive_p+belowfive_np)*100, digits = 2), round(five_fifty_p/(five_fifty_p+five_fifty_np)*100, digits = 2), round(fifty_hund_p/(fifty_hund_p+fifty_hund_np)*100, digits = 2), round(hund_fifty_p/(hund_fifty_p+hund_fifty_np)*100, digits = 2), round(fifty_twohund_p/(fifty_twohund_p+fifty_twohund_np)*100, digits = 2), round(twohund_fifty_p/(twohund_fifty_p+twohund_fifty_np)*100, digits = 2), round(above_p/(above_p+above_np)*100, digits = 2)),
+  Not_in_poverty_percentage = c(round(belowfive_np/(belowfive_p+belowfive_np)*100, digits = 2), round(five_fifty_np/(five_fifty_p+five_fifty_np)*100, digits = 2), round(fifty_hund_np/(fifty_hund_p+fifty_hund_np)*100, digits = 2), round(hund_fifty_np/(hund_fifty_p+hund_fifty_np)*100, digits = 2), round(fifty_twohund_np/(fifty_twohund_p+fifty_twohund_np)*100, digits = 2), round(twohund_fifty_np/(twohund_fifty_p+twohund_fifty_np)*100, digits = 2), round(above_np/(above_p+above_np)*100, digits = 2))
+)
+
+kable(income_level, 
+      align = c("c", "c", "c")  # Center alignment for all columns
+)
+```
+
+```{r fig.show='hide'}
 #| label: fig-income
@@
 
-\newpage
-
 # Model {#sec-model}
-In our analysis, we utilized a Bayesian logistic regression model to examine the relationship between poverty status and two demographic factors, mortgage status and income of the household.
+In our analysis, we utilized a Bayesian logistic regression model to examine the relationship between poverty status and two financial factors, mortgage status and income of the household.
 Background details and diagnostics are included in [Appendix -@sec-model-details].
@@
 
-\newpage
 # Discussion {#sec-discussion}
@@
 The percentage of individuals below the poverty threshold, adjusted for housing costs are: $29,080 for a family of four with a mortgage, $24,413 for a family of four without a mortgage, and $29,194 for a family of four renting [@Khattar_Pathak_Schweitzer_Khan_McManus_Jones_Benson_Kelly_2023].
-From this, we can assumed that families without mortgages experienced the lowest housing costs, leaving more funds available for other essentials.
-In contrast, families with mortgages may face burdensome monthly payments, especially if they have chosen a floating interest rate, making them vulnerable to fluctuations in interest rates.
-Renters also encounter challenges, as rent increases can strain finances and necessitate either accepting higher housing costs or undertaking the time-consuming and costly process of finding new accommodations.
+From this, we can see that families without mortgages experienced the lowest housing costs, leaving more funds available for other essentials.
+In contrast, families with mortgages may face expensive monthly payments, especially if they have chosen a floating interest rate, making them vulnerable to rising interest rates.
+Renters also experience challenges when the rent increases.
+So they either have to accept higher housing costs or waste time and energy to find new homes.
+This supports my model result that renters are more likely to be in poverty.
 
-Also, low-income households face challenges with accessing lower mortgage rates and reducing debt burdens through refinancing.
-Households with low to moderate incomes were less able to take advantage of lower mortgage rates and alleviate their debt burdens through refinancing during the period of exceptionally low rates: merely 24% of mortgages in low-income regions underwent refinancing, in contrast to the 42% in high-income areas [@mortgage].
-Therefore, the low-income households have harder experience with repaying their mortgage than higher income households.
+Also, low-income households have a harder time to access lower mortgage rates and reduce debt burdens through refinancing.
+Households with low to moderate incomes were less able to take advantage of lower mortgage rates and relieve their debt burdens through refinancing during the period of exceptionally low rates: only 24% of mortgages in low-income regions underwent refinancing, in contrast to the 42% in high-income areas [@mortgage].
+Therefore, it is generally harder for low-income households to repay their mortgage than higher income households.
 
-These housing-related challenges increases the risk of poverty for affected families, underscoring the importance of addressing housing affordability as a key component of poverty alleviation efforts.
+These problems with housing make it more likely for families which are already in poverty, to become poorer.
+So it is important to make housing more affordable for people, in order to help people out of poverty.
 
@@
 However, in the United States, the cost of living is notably high, low-income households often struggle to meet basic needs, resulting in an unending cycle of poverty.
-About half of lower-income parents (52%) say they have not had enough money for food or their rent or mortgage [@Braga_2022].
-Some 45% of employed parents with lower incomes say they would be very worried about losing pay if they needed to take a day or two off from work to deal with a sick child or because of child care issues.
+About half of lower-income parents (52%) say they have not had enough money for food or their rent or mortgage.
+Some 45% of employed parents with lower incomes say they would be very worried about losing pay if they needed to take a day or two off from work to deal with a sick child or because of child care issues [@Braga_2022].
 This type of delimma for low-income households makes them harder to pull themselves out of poverty.
@@
 So, it is evident that families earning less than $5,000 are definitely living below the poverty line.
+This is supported by my model result.
 Conversely, individuals with high incomes are less likely to experience poverty, as their financial resources enable them to comfortably cover living expenses.
@@
 
+## US Government Programs to Help
 For household in poverty, escaping poverty is tough, but the US government and communities are stepping up to help.
@@
 
-### Data Consideration
-Firstly, the data are estimates based on a sample of the population, so it cannot represent the whole population of the United States.
-This creates margins of error.
-Also, since the SPM poverty estimates are from household survey data, the data might be affected by issues such as under-reporting, over-reporting, or making up data.
-Some people might not remember the exact value of some data, for example, not all people have or know that they receive income from government benefit programs, so they might be reporting it inaccurately.
-In addition, some SPM resources, such as taxes and non-cash benefits, are estimated using models.
-The models use administrative data, but the estimated amounts are not perfect.
-
 ### Measurement of Medical Subsidies
 Even the SPM accounts for expenses for healthcare and insurance by subtracting them from family resources, it does not directly value health insurance provided as a resource [@spm].
-The treatment of medical expenses does not take account of all the economic effects of subsidized medical care, for example, there are families who purchase a less comprehensive insurance plan and decide to forego some health care.
+The treatment of medical expenses does not take account of all the economic effects of subsidized medical care, for example, there are families who purchase a less expensive insurance plan and decide to give up some health care.
 Moreover, some of the largest non-cash benefits programs (Medicare, Medicaid, etc.) are not explicitly taken into account when determining poverty status.

### 2024-04-18 paper/US_Poverty_Analysis_2019.qmd
final checks

@@
 Federal Statistical System, responsible for producing data about the American people and economy [@uscb].
-Over the years, people have realized that the Official Poverty Measure is flawed in various aspects, such as not accounting for regional differences in costs of living, not considering household's mortgage state, and government benefits programs  [@Gesumaria_2015].
+Over the years, people have realized that the Official Poverty Measure is flawed in various aspects [@Gesumaria_2015].
 Therefore, Supplemental Poverty Measure was developed to address numerous concerns of Official Poverty Measure critics, providing an improved statistical picture of poverty.

### 2024-04-18 paper/US_Poverty_Analysis_2019.qmd
graph and table

@@
 ## Data Visualization
-@tbl-mortgage illustrates the relationship between mortgage state and poverty status of households.
+@tbl-mortgage and @tbl-mortgage illustrates the relationship between mortgage state and poverty status of households.
 It can be seen that renters are the highest proportion of households experiencing poverty, followed by owners without mortgages, and finally owners with mortgages.
@@
 
-```{r fig.show='hide'}
+```{r}
 #| label: fig-mortgage
@@
 
-@tbl-income shows the relationship between household annual total income and their poverty status.
+@fig-income and @tbl-income shows the relationship between household annual total income and their poverty status.
 We can see that among households earning more than 100k annually, no household is in poverty.
@@
 
-```{r fig.show='hide'}
+```{r}
 #| label: fig-income

### 2024-04-18 paper/US_Poverty_Analysis_2019.qmd
final edits

@@
 ## Data Visualization
-@tbl-mortgage and @tbl-mortgage illustrates the relationship between mortgage state and poverty status of households.
+@fig-mortgage and @tbl-mortgage(in Appendix @sec-data-details) illustrate the relationship between mortgage state and poverty status of households.
 It can be seen that renters are the highest proportion of households experiencing poverty, followed by owners without mortgages, and finally owners with mortgages.
@@
 Despite variations in poverty rates across different mortgage states, the cumulative data shows economic stability among the majority of households in the United States in 2019.
-```{r}
-#| label: tbl-mortgage
-#| tbl-cap: "The distribution of poverty status by mortgage state"
-#| message: false
-#| echo: false
-#| warning: false
-colnames(poverty2019) <- c("poverty_status", "mortgage_state", "income")
-
-# calculate poverty ratio for renter, Owner with Mortgage, and Owner without Mortgage
-rent_p <- sum(poverty2019$mortgage_state == "Renter" & poverty2019$poverty_status == "In poverty")
-rent_np <- sum(poverty2019$mortgage_state == "Renter" & poverty2019$poverty_status == "Not in poverty")
-
-mort_p <- sum(poverty2019$mortgage_state == "Owner with Mortgage" & poverty2019$poverty_status == "In poverty")
-mort_np <- sum(poverty2019$mortgage_state == "Owner with Mortgage" & poverty2019$poverty_status == "Not in poverty")
-
-nomort_p <- sum(poverty2019$mortgage_state == "Owner without Mortgage" & poverty2019$poverty_status == "In poverty")
-nomort_np <- sum(poverty2019$mortgage_state == "Owner without Mortgage" & poverty2019$poverty_status == "Not in poverty")
-
-# show data feature
-mortgage <- tibble(
-  Household_type = c("Renters", "Owners with Mortgage", "Owners without Mortgage"),
-  In_poverty_percentage = c(round(rent_p/(rent_p+rent_np)*100, digits = 2), round(mort_p/(mort_p+mort_np)*100, digits = 2), round(nomort_p/(nomort_p+nomort_np)*100, digits = 2)),
-  Not_in_poverty_percentage = c(round(rent_np/(rent_p+rent_np)*100, digits = 2), round(mort_np/(mort_p+mort_np)*100, digits = 2), round(nomort_np/(nomort_p+nomort_np)*100, digits = 2))
-)
-
-kable(mortgage, 
-      align = c("c", "c", "c")  # Center alignment for all columns
-)
-```
 
@@
 
-@fig-income and @tbl-income shows the relationship between household annual total income and their poverty status.
+@fig-income and @tbl-income(in Appendix @sec-data-details) show the relationship between household annual total income and their poverty status.
 We can see that among households earning more than 100k annually, no household is in poverty.
@@
 This indicates a strong positive correlation between higher income levels and financial stability.
-
-```{r}
-#| label: tbl-income
-#| tbl-cap: "The distribution of poverty status by income level"
-#| message: false
-#| echo: false
-#| warning: false
-
-# calculate poverty ratio for different income levels
-five_fifty_p <- sum(poverty2019$income == "5k-50k" & poverty2019$poverty_status == "In poverty")
-five_fifty_np <- sum(poverty2019$income == "5k-50k" & poverty2019$poverty_status == "Not in poverty")
-
-belowfive_p <- sum(poverty2019$income == "below 5k" & poverty2019$poverty_status == "In poverty")
-belowfive_np <- sum(poverty2019$income == "below 5k" & poverty2019$poverty_status == "Not in poverty")
-
-fifty_hund_p <- sum(poverty2019$income == "50k-100k" & poverty2019$poverty_status == "In poverty")
-fifty_hund_np <- sum(poverty2019$income == "50k-100k" & poverty2019$poverty_status == "Not in poverty")
-
-hund_fifty_p <- sum(poverty2019$income == "100k-150k" & poverty2019$poverty_status == "In poverty")
-hund_fifty_np <- sum(poverty2019$income == "100k-150k" & poverty2019$poverty_status == "Not in poverty")
-
-fifty_twohund_p <- sum(poverty2019$income == "150k-200k" & poverty2019$poverty_status == "In poverty")
-fifty_twohund_np <- sum(poverty2019$income == "150k-200k" & poverty2019$poverty_status == "Not in poverty")
-
-twohund_fifty_p <- sum(poverty2019$income == "200k-250k" & poverty2019$poverty_status == "In poverty")
-twohund_fifty_np <- sum(poverty2019$income == "200k-250k" & poverty2019$poverty_status == "Not in poverty")
-
-above_p <- sum(poverty2019$income == "above 250k" & poverty2019$poverty_status == "In poverty")
-above_np <- sum(poverty2019$income == "above 250k" & poverty2019$poverty_status == "Not in poverty")
-
-# show data feature
-income_level <- tibble(
-  Household_type = c("below 5k","5k-50k","50k-100k","100k-150k","150k-200k","200k-250k","above 250k"),
-  In_poverty_percentage = c(round(belowfive_p/(belowfive_p+belowfive_np)*100, digits = 2), round(five_fifty_p/(five_fifty_p+five_fifty_np)*100, digits = 2), round(fifty_hund_p/(fifty_hund_p+fifty_hund_np)*100, digits = 2), round(hund_fifty_p/(hund_fifty_p+hund_fifty_np)*100, digits = 2), round(fifty_twohund_p/(fifty_twohund_p+fifty_twohund_np)*100, digits = 2), round(twohund_fifty_p/(twohund_fifty_p+twohund_fifty_np)*100, digits = 2), round(above_p/(above_p+above_np)*100, digits = 2)),
-  Not_in_poverty_percentage = c(round(belowfive_np/(belowfive_p+belowfive_np)*100, digits = 2), round(five_fifty_np/(five_fifty_p+five_fifty_np)*100, digits = 2), round(fifty_hund_np/(fifty_hund_p+fifty_hund_np)*100, digits = 2), round(hund_fifty_np/(hund_fifty_p+hund_fifty_np)*100, digits = 2), round(fifty_twohund_np/(fifty_twohund_p+fifty_twohund_np)*100, digits = 2), round(twohund_fifty_np/(twohund_fifty_p+twohund_fifty_np)*100, digits = 2), round(above_np/(above_p+above_np)*100, digits = 2))
-)
-
-kable(income_level, 
-      align = c("c", "c", "c")  # Center alignment for all columns
-)
-```
 
@@
 ```
+
+## Data Visualization
+@tbl-mortgage is the percentage distribution of poverty status by mortgage state.
+```{r}
+#| label: tbl-mortgage
+#| tbl-cap: "The percentage distribution of poverty status by mortgage state"
+#| message: false
+#| echo: false
+#| warning: false
+colnames(poverty2019) <- c("poverty_status", "mortgage_state", "income")
+
+# calculate poverty ratio for renter, Owner with Mortgage, and Owner without Mortgage
+rent_p <- sum(poverty2019$mortgage_state == "Renter" & poverty2019$poverty_status == "In poverty")
+rent_np <- sum(poverty2019$mortgage_state == "Renter" & poverty2019$poverty_status == "Not in poverty")
+
+mort_p <- sum(poverty2019$mortgage_state == "Owner with Mortgage" & poverty2019$poverty_status == "In poverty")
+mort_np <- sum(poverty2019$mortgage_state == "Owner with Mortgage" & poverty2019$poverty_status == "Not in poverty")
+
+nomort_p <- sum(poverty2019$mortgage_state == "Owner without Mortgage" & poverty2019$poverty_status == "In poverty")
+nomort_np <- sum(poverty2019$mortgage_state == "Owner without Mortgage" & poverty2019$poverty_status == "Not in poverty")
+
+# show data feature
+mortgage <- tibble(
+  Household_type = c("Renters", "Owners with Mortgage", "Owners without Mortgage"),
+  In_poverty_percentage = c(round(rent_p/(rent_p+rent_np)*100, digits = 2), round(mort_p/(mort_p+mort_np)*100, digits = 2), round(nomort_p/(nomort_p+nomort_np)*100, digits = 2)),
+  Not_in_poverty_percentage = c(round(rent_np/(rent_p+rent_np)*100, digits = 2), round(mort_np/(mort_p+mort_np)*100, digits = 2), round(nomort_np/(nomort_p+nomort_np)*100, digits = 2))
+)
+
+kable(mortgage, 
+      align = c("c", "c", "c")  # Center alignment for all columns
+)
+```
+
+@tbl-income is the percentage distribution of poverty status by income level.
+```{r}
+#| label: tbl-income
+#| tbl-cap: "The percentage distribution of poverty status by income level"
+#| message: false
+#| echo: false
+#| warning: false
+
+# calculate poverty ratio for different income levels
+five_fifty_p <- sum(poverty2019$income == "5k-50k" & poverty2019$poverty_status == "In poverty")
+five_fifty_np <- sum(poverty2019$income == "5k-50k" & poverty2019$poverty_status == "Not in poverty")
+
+belowfive_p <- sum(poverty2019$income == "below 5k" & poverty2019$poverty_status == "In poverty")
+belowfive_np <- sum(poverty2019$income == "below 5k" & poverty2019$poverty_status == "Not in poverty")
+
+fifty_hund_p <- sum(poverty2019$income == "50k-100k" & poverty2019$poverty_status == "In poverty")
+fifty_hund_np <- sum(poverty2019$income == "50k-100k" & poverty2019$poverty_status == "Not in poverty")
+
+hund_fifty_p <- sum(poverty2019$income == "100k-150k" & poverty2019$poverty_status == "In poverty")
+hund_fifty_np <- sum(poverty2019$income == "100k-150k" & poverty2019$poverty_status == "Not in poverty")
+
+fifty_twohund_p <- sum(poverty2019$income == "150k-200k" & poverty2019$poverty_status == "In poverty")
+fifty_twohund_np <- sum(poverty2019$income == "150k-200k" & poverty2019$poverty_status == "Not in poverty")
+
+twohund_fifty_p <- sum(poverty2019$income == "200k-250k" & poverty2019$poverty_status == "In poverty")
+twohund_fifty_np <- sum(poverty2019$income == "200k-250k" & poverty2019$poverty_status == "Not in poverty")
+
+above_p <- sum(poverty2019$income == "above 250k" & poverty2019$poverty_status == "In poverty")
+above_np <- sum(poverty2019$income == "above 250k" & poverty2019$poverty_status == "Not in poverty")
+
+# show data feature
+income_level <- tibble(
+  Household_type = c("below 5k","5k-50k","50k-100k","100k-150k","150k-200k","200k-250k","above 250k"),
+  In_poverty_percentage = c(round(belowfive_p/(belowfive_p+belowfive_np)*100, digits = 2), round(five_fifty_p/(five_fifty_p+five_fifty_np)*100, digits = 2), round(fifty_hund_p/(fifty_hund_p+fifty_hund_np)*100, digits = 2), round(hund_fifty_p/(hund_fifty_p+hund_fifty_np)*100, digits = 2), round(fifty_twohund_p/(fifty_twohund_p+fifty_twohund_np)*100, digits = 2), round(twohund_fifty_p/(twohund_fifty_p+twohund_fifty_np)*100, digits = 2), round(above_p/(above_p+above_np)*100, digits = 2)),
+  Not_in_poverty_percentage = c(round(belowfive_np/(belowfive_p+belowfive_np)*100, digits = 2), round(five_fifty_np/(five_fifty_p+five_fifty_np)*100, digits = 2), round(fifty_hund_np/(fifty_hund_p+fifty_hund_np)*100, digits = 2), round(hund_fifty_np/(hund_fifty_p+hund_fifty_np)*100, digits = 2), round(fifty_twohund_np/(fifty_twohund_p+fifty_twohund_np)*100, digits = 2), round(twohund_fifty_np/(twohund_fifty_p+twohund_fifty_np)*100, digits = 2), round(above_np/(above_p+above_np)*100, digits = 2))
+)
+
+kable(income_level, 
+      align = c("c", "c", "c")  # Center alignment for all columns
+)
+```
+
 # Model details {#sec-model-details}

### 2024-04-19 scripts/01-download_data.R
some edits

@@ -11,6 +11,7 @@
 #### Workspace setup ####
 library(haven)
 library(tidyverse)
+library(arrow)
 
 
 #### Download data ####
@@ -23,6 +24,6 @@ raw_poverty_data <-
 
 
 #### Save data ####
-write_csv(raw_poverty_data, "data/raw_data/raw_poverty_data.csv") 
+write_parquet(raw_poverty_data, sink = "data/raw_data/raw_poverty_data.parquet")
 
          

### 2024-04-19 paper/US_Poverty_Analysis_2019.qmd
final checks

@@
 # read in raw data
-raw_poverty_data <-  read_csv(here::here("data/raw_data/raw_poverty_data.csv"))
+raw_poverty_data <-  read_parquet(here::here("data/raw_data/raw_poverty_data.parquet"))
 ```
@@
 Regarding the relationship between mortgage state and poverty state, we anticipate that both owners without mortgages and owners with mortgages are less likely to experience poverty.
-Owners without mortgages typically indicate financial stability, as they have already paid off their homes and can allocate financial resources toward household necessities, other than paying back the mortgage / rent.
+Owners without mortgages typically indicate financial stability, as they have already paid off their homes and can allocate financial resources toward household necessities.
 Also, without monthly mortgage or rent payments, their financial burden is reduced, further decreasing the likelihood of poverty.
@@
 
-In terms of income levels, we expect a positive relationship between household's income level and their poverty status.
-This expectation arises from common sense that higher income typically reduces the likelihood of experiencing poverty.
+In terms of income levels, we expect a negative relationship between household's income level and their likelihood to be in poverty.
+This expectation arises from our common sense that higher income typically reduces the likelihood of experiencing poverty.
 If a family is earning a lot of money, then naturally it would not be in poverty.
@@
 This correlation shows the role of financial resources in mitigating economic hardship and facilitating socio-economic well-being.
-However, in the United States, the cost of living is notably high, low-income households often struggle to meet basic needs, resulting in an unending cycle of poverty.
-About half of lower-income parents (52%) say they have not had enough money for food or their rent or mortgage.
-Some 45% of employed parents with lower incomes say they would be very worried about losing pay if they needed to take a day or two off from work to deal with a sick child or because of child care issues [@Braga_2022].
-This type of delimma for low-income households makes them harder to pull themselves out of poverty.
-
 In 2019, the poverty guideline for a one-person household was set at $12,490 annually [@ASPE].
@@
 This is supported by my model result.
-Conversely, individuals with high incomes are less likely to experience poverty, as their financial resources enable them to comfortably cover living expenses.
+Individuals with high incomes are less likely to experience poverty, as their financial resources enable them to comfortably cover living expenses.
 They will be able to have savings accumulation, further consolidating their financial security and contributing to wealth accumulation over time.
+
+In the United States, the cost of living is notably high, low-income households often struggle to meet basic needs, resulting in an unending cycle of poverty.
+About half of lower-income parents (52%) say they have not had enough money for food or their rent or mortgage.
+45% of employed parents with lower incomes say they would be very worried about losing pay if they needed to take a day or two off from work to deal with a sick child or because of child care issues [@Braga_2022].
+This type of delimma for low-income households makes them harder to pull themselves out of poverty.
 
@@
 Programs like the Child Tax Credit (CTC) [@ctc], which provides tax credits to eligible families with qualifying children.
+The credit currently provides up to $2,000 per child to about 40 million families every year [].
+So obviously, these programs are a huge help for low-income family with kids.
 Other programs, like housing assistance and Medicaid, provide support with things like affordable housing and healthcare.

### 2024-04-19 scripts/02-data_cleaning.R

@@ -15,7 +15,7 @@ library(arrow)
 #### Clean data ####
 # read in raw data
 raw_poverty_data <-
-  read_csv("data/raw_data/raw_poverty_data.csv")
+  read_parquet("data/raw_data/raw_poverty_data.parquet")
 
 # Calculate money income after tax, minus expenses, plus credits and benefits
 raw_poverty_data$income_after_tax <- raw_poverty_data$spm_totval + raw_poverty_data$spm_snapsub + raw_poverty_data$spm_caphousesub + raw_poverty_data$spm_schlunch + raw_poverty_data$spm_engval + raw_poverty_data$spm_wicval - raw_poverty_data$spm_fedtax + raw_poverty_data$spm_eitc + raw_poverty_data$spm_actc - raw_poverty_data$spm_fica - raw_poverty_data$spm_sttax - raw_poverty_data$spm_childsuppd - raw_poverty_data$spm_capwkccxpns - raw_poverty_data$spm_medxpns

### 2024-04-19 paper/datasheet_2019SPM.qmd
datasheet edit

@@
 date-format: long
+abstract: "Provides access to 2019 Revised Supplemental Poverty Measure Research Dataset about American households' demographic and financial information.
+This enhances how reproducible research is that uses this data."
 format: pdf

### 2024-04-19 paper/US_Poverty_Analysis_2019.qmd
final edits

@@
 
-The logistic regression model shows that in the US in 2019, households with an annual total income less than $50000 are almost all in poverty.
-For households earning $50000 to $100000, the majority is not in poverty, yet a notable amount of them are still in poverty.
+The logistic regression model shows that in the US in 2019, households with an annual total income less than $5000 are in poverty.
+For households earning $5000 to $50000, the majority is not in poverty, yet some of them are still in poverty.
 For households earning above $100000, all of them are not in poverty.
@@
 In the USA in 2019, housing costs accounted for more than 30% of household expenditure, making mortgage status a critical factor in determining a family's risk of falling into poverty [@Statistics_2020].
-The percentage of individuals below the poverty threshold, adjusted for housing costs are: $29,080 for a family of four with a mortgage, $24,413 for a family of four without a mortgage, and $29,194 for a family of four renting [@Khattar_Pathak_Schweitzer_Khan_McManus_Jones_Benson_Kelly_2023].
+A study shows that in 2019, the annual income needed are: $29,080 for a family of four with a mortgage; $24,413 for a family of four without a mortgage; and $29,194 for a family of four renting [@Khattar_Pathak_Schweitzer_Khan_McManus_Jones_Benson_Kelly_2023].
 From this, we can see that families without mortgages experienced the lowest housing costs, leaving more funds available for other essentials.
@@
 So they either have to accept higher housing costs or waste time and energy to find new homes.
-This supports my model result that renters are more likely to be in poverty.
+This supports my model result that renters are more likely to be in poverty, followed by owners with mortgage, then owners without mortgage.
 
@@
 Therefore, it is generally harder for low-income households to repay their mortgage than higher income households.
-
-These problems with housing make it more likely for families which are already in poverty, to become poorer.
+The challenge of paying mortgage might increase household's likelihood to be in poverty.
+We can see that these problems with housing make it more likely for families which are already in poverty, to become poorer.
 So it is important to make housing more affordable for people, in order to help people out of poverty.
@@
 Programs like the Child Tax Credit (CTC) [@ctc], which provides tax credits to eligible families with qualifying children.
-The credit currently provides up to $2,000 per child to about 40 million families every year [].
+The credit currently provides up to $2,000 per child to about 40 million families every year [@child].
 So obviously, these programs are a huge help for low-income family with kids.
@@
 ### Lump sum nature of tax credit
-In the SPM, taxes paid and tax credits in the year that are earned are include [@spm]d.
+In the SPM, taxes paid and tax credits in the year that are earned are included [@spm].
 But typically the credit would not appear in the family income until the next year.
